---
title: Benefits and Costs of Dual and Informal Apprenticeship in Bénin
author: 
  - name: Bart Kudrzycki
    affiliation: Development Economics Group, ETH Zurich, Switzerland, <bartlomiej.kudrzycki@nadel.ethz.ch>
header-includes:
  - \usepackage{setspace}
date: \today
abstract: |

  Traditional apprenticeships are an important source of skills for early school leavers in developing countries. Usually private arrangements between parents and informal firms, apprenticeship quality and costs are not subject to oversight or regulation. Public reforms have aimed at improving access or quality of training, for example by complementing in-firm training with a weekly classroom component. We provide a detailed account of such a program conducted at national scale in Bénin. Using two waves of surveys with firm owners and apprentices, we analyze the human capital gains and material costs and benefits associated with training. We find no differences in human capital accumulation between apprentices who participated in the program to a sample of rejected applicants and non-applicants from the same firms. Firms which trained more dual apprentices did not grow more quickly or become more profitable. Additionally, we observe that allowances distributed to apprentices in lieu of wages appear to be considerably higher than training fees paid, especially at larger firms. This contradicts the common wisdom that informal firms rely on apprenticeship fees as a source of financing. Instead, apprentice productive contributions to the firm appear to be the primary incentive for firms to participate in training. **The dual training program needs improvement, not clear why more experienced (hence productive) apprentices are sent; balance must be struck between the effectiveness of classroom training (should be weekly) and the preferences of trainers, who apparently benefit from apprentice labor, not fees; as argued by others, increasing MC skin in the game may be key for masters to see the classroom component as a benefit.**

# This paper evaluates the impact of and provides a detailed account of the costs and benefits associated with (dual) apprenticeship training in the informal sector in Bénin using unique matched apprentice-firm survey data. To study the effectiveness of dual training in an informal economy, successful applicants to the _Certificat de Qualification Professionelle_, program, a national training program that combines in-firm training with classroom teaching, are compared to unsuccessful applicants and non-applicants three years after program start.  Human capital gains among all apprentices, as measured by trainers' subjective appraisal of apprentice experience and competence, are significant, with improvements of .73 and .66 standard deviation between survey waves, respectively. Gains were 3concentrated in the masonry and carpentry sector, in larger firms, and among apprentices who did not apply for the CQP program. To study net benefits of training for firms, two cost and benefit models are used. When only fees, allowances, and reported training costs are considered, firms suffer a net loss of 130 $US per apprentice per year: an average of 630 $US total per firm. When projected apprentice contributions and foregone trainer productivity are included, net estimated losses are reduced to 33 \$US per apprentice per year, and an overall estimated benefit of 218 $US per firm. Fees paid to the training firms account are not enough to defray the expenses associated with training, which are dominated by allowances paid to apprentices. Apprentices who do not apply to the CQP program are associated with somewhat lower net losses to their training firms when estimated using the first model. Apprentices hired at larger firms are associated with higher net benefits. In sum, we find that participation in the CQP program is not associated in any benefits for either trainees or firms relative to non-participating apprentices, though effectiveness of dual training may have been negatively impacted by the Covid-19 crisis.
  
thanks: |
  Thanks to Dario for the Markdown template.
keywords: [Informal labor markets, Dual training, Apprenticeship]
jel: [I26]
bibliography: cqp-bib.bib
csl: 'apa-no-ampersand.csl'
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    template: template.tex
    extra_dependencies: ["flafter", "placeins"]
---

```{r setup, include=FALSE}
packages <- c("tinytex", "lfe", "lme4", "plm", "bookdown", "gtsummary", "labelled", "grid", "gridExtra", "rstatix", "kableExtra", "stargazer", "tidyverse")

options(repos = c(CRAN = "http://cran.rstudio.com"))

# install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages], silent = TRUE)
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages, installed_packages)

# load functions
source("../functions/add_by_n.R")

# load data
# source("../code/master.R") # to rerun all code periodically
load("../data/df.rda")
selected <- df %>% filter(SELECTED == 1)
baseline <- df %>% filter(wave == 0)
endline <- df %>% filter (wave == 1)
firms <- df %>% select(FS1.2, wave, FS3.4, firm_size, FS6.1, contains("cb")) %>% group_by(FS1.2, wave) %>% summarise_all(mean, na.rm = T) %>% ungroup()

# load code chunks
knitr::read_chunk('../code/tbls_body.R')
knitr::read_chunk('../code/tbls_appendix.R')
knitr::read_chunk('../code/figures.R')

# set R options
ggplot2::theme_set(ggplot2::theme_bw())
options(scipen=2, digits=2)

# set knitr options
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      echo = FALSE,
                      cache = TRUE,
                      fig.path='figures/',
                      out.extra = "",
                      floatfoot.envs = c("figure", "table", "longtabu"))
```


\doublespacing

# Introduction {#intro}

In sub-Saharan Africa (SSA), interest in apprenticeships is on the rise. In economies with largely informal economies, traditional apprenticeships (also referred to as apprenticeships in the informal sector or informal apprenticeships) are the most important source of skills for early school leavers, accounting for as much as 80 percent of technical and vocational training (TVET) in SSA [@filmer2014], and for as much as 90 percent of total employment in the crafts sector [@walther2007; @worldbank2017]. According to a survey of five countries, 20 percent of youth aged 25-34 had participated in apprenticeship; in contrast, only 1 percent of youth participate in formal TVET, and about 9 percent are enrolled in tertiary education [@filmer2014]. For youth who drop out of school early, informal on-the-job training is often the only alternative for acquiring the skills necessary to start a business or find employment. As increasing numbers of youth in SSA suffer from a lack of labor attachment, underemployment, and poverty, informal sector training is seen by many policy experts as an important tool to tackle the youth employment challenge [@filmer2014].

Traditional apprenticeships take place predominantly in the informal sector and generally last between three and four years. They involve a private contractual arrangement between an apprentice - usually a school-leaver between the ages of 14 and 18 - or his or her parents and a master craftsman (MC) who trains the apprentice in the workplace for a fee [@bas1989]. In contrast to formal TVET, which takes place almost exclusively in the classroom in the SSA context, traditional apprentices train on-the-job. Upon completion of the apprenticeship, the MC issues a certificate acknowledging the training; some apprentices continue to work for the same or for another workshop as a wage employee, though most seek to start their own firm given access to sufficient capital [@frazer2006]. While unregulated at the national level, informal apprenticeships are nevertheless structured according to the dictates of tradition and the customs of local professional associations, and, in the context of highly informal economies, are generally considered to be more effective than formal TVET at delivering the skills demanded by the labor market [@ahadzie2009].

The unregulated nature of traditional apprenticeships give rise to a number of potential market failures that negatively affect their provision, and have led to calls for their reform [@walther2011]. For instance, in the absence of complete, enforceable contracts, firms may be unable to commit to providing general skills training [@acemogluWhyFirmsTrain1998; @acemogluStructureWagesInvestment1999; @dustmann2012]. Apprentice productivity may also be so low that subsistence levels (paid in the form of "chop money" by the firm owner) are greater returns from training, causing an under-provision of training. Another obstacle to training provision may involve fears of competing firms "poaching" newly trained apprentices, a problem particularly salient for small enterprises, which experience higher employee turnover and offer fewer opportunities for career advancement [@mcintosh2011], though evidence of this is limited for the African context. Quality is also affected by the unregulated nature of informal apprenticeships, and may have adverse consequences for participating youth. For instance, apprentices may be exposed to inexperienced trainers who keep them in their apprenticeship for too long [@bas1989], or experience limited labor market mobility into formal sector wage jobs due to the lack of formal accreditation systems [@acemogluCertificationTrainingTraining2000; @worldbank2017; @alfonsi2020].

A deeper understanding of the costs and benefits are critical for guiding reforms to address such market failures. Policies introducing competence-based, nationally-accredited certification of informal apprenticeship are thus a popular solution to this issue, and have been recently introduced in SSA in countries including Malawi and Tanzania.

Another proposed reform is the introduction of a classroom component to traditional apprenticeship, producing a hybrid "dual system" comparable to the Swiss and German variety [@walther2011]. Dual systems promise to increase training quality by introducing a state-regulated classroom component, while also improving the signalling ability of apprentices upon graduation by offering official certification. In SSA, dual apprenticeship schemes were first introduced in Bénin and Togo in the 1980s by the Hans Seidel Foundation, a German NGO, and apprenticeship reforms based on the dual system have since been introduced in Mali, Côte d'Ivoire, Senegal, Tanzania, Togo, and Niger [@walther2011; @ilo2020a]. Many of these schemes have struggled with funding issues and integration into the existing national TVET and regulatory frameworks. With their potential to simultaneously harness the abundance of training firms in the informal sector and the growing demand among parents and youth for formal education, dual system apprenticeship certification schemes seem to remain an underutilized approach in the field of TVET reform.

In this paper, we employ apprentice-firm survey data collected in Bénin in two waves (baseline in 2019 and endline in 2021) for 427 apprentices training in 197 firms to analyze the impact of a national dual system training program on participating apprentices and firms. The program offered youth in select trades the opportunity to attend classroom training at a local training center once a week while continuing with their traditional in-firm apprenticeship training with an MC. Youth and firms were interviewed before the onset of training and after three years (somewhat shorter than the typical duration of an apprenticeship). We use a sector-specific skills test to compare the learning outcomes of dual system apprentices to unsuccessful program applicants and non-applicants (with both comparison groups pursuing traditional apprenticeships at the same firms as dual-system participants). We also estimate the marginal effect of apprentice participation in dual system training on firm size and profits.

Studies of vocational training interventions combining on-the-job and classroom teaching in middle-income countries have reported persistent if modest increases in earnings and mixed impacts on employment [@card2011; @attanasio2011; @ibarraran2014; @alzua2016;  @attanasio2017; @ibarraran2019], though similar interventions in LICs have been characterized by low take-up, high dropout, and low efficacy [see @blattman2015; @ghisletta2021; @tripney2013 for an overview]. These programs tend to be shorter (several months, rather than years) than the one studied in this paper, and focus on employment in the formal rather than the informal sector. To our knowledge, only one paper has addressed dual-system training in Sub-Saharan Africa. In a randomized experiment in Côte d'Ivoire, @crepon2019 find that youth offered a stipend for an apprenticeship that combined 12 to 24 months of on-the-job training with theoretical classes at local training institutions earn 15 percent more after three years, partake in more complex and non-routine tasks\footnote{The authors find that treated youth who enroll in the subsidized dual training program are more likely to be involved in complex, non-routine tasks after two years, with the likelihood to undertake non-routine analytical tasks increasing by .24 standard deviations (SDs) and non-routine interpersonal tasks increasing by 0.08 SDs relative to non-treated traditional apprentices. A task intesity index was found to be .21 SDs lower for dual apprentices, suggesting that dual apprentices are involved in a wider range of tasks.}, and receive training certification at a higher rate than non-treated youth. We study a similarly structured program, but one that is about twice as long and does not involve any direct subsidies or eliminate fees. Our study also contributes quantitative evidence regarding the effectiveness of the CQP to the discussion tracing the program's history and current implementation [@davodoun2011; @david-gnahoui2017; @bankole2020].

A related literature addresses the profitability of dual-system apprenticeships for the employer through the detailed accounting of the costs and benefits of training. A number of such studies have been conducted in the European context with the aid of surveys and simulations [see, e.g., @muhlemann2016; @muhlemann2019; @muhlemann2014], but have only recently begun to generate interest in lower-middle and low-income countries. Examples include @bolli2021, who find that training costs outweigh benefits in Serbia (with larger firms suffering smaller losses), and @bolli2020, who show that training firms in Nepal generally profit from training, with little variation across firm size. To our knowledge, ours is the first rigorous cost-benefit study of dual-system training conducted in SSA.

We also contribute to the literature on informal apprenticeship in SSA as such by considering effect of training on both apprentices and firms. Evidence from high-income countries suggests that completing an apprenticeship improves overall labor market outcomes for participating youth, both in terms of employment and earnings [@oecdinternationallabourorganization2017]. Though traditional apprenticeships are very common in West Africa [@adams2013; @filmer2014; @walther2011], there is limited direct empirical evidence on their impact on the labor market outcomes of apprentices. Long-term returns to informal training have been shown to be quite heterogenous in Ghana, while benefiting youth with lower levels of education the most [@monk2008]. An experimental study in Uganda finds that six months of in-firm training measurably improves skills, and that these skills persist two to three years after the end of training [@alfonsi2020]. However, skills acquired in informal training tend to be firm-specific, and thus more likely to lead to self-employment than to quick career progression in the formal sector than formal schooling [@frazer2006; @hardy2019; @alfonsi2020]. 

Studies from SSA suggest that informal apprenticeship training may increase microenterprise size and profitability. Using data on formal manufacturing firms from Kenya, Zimbabwe and Ghana, @rosholm2007 observe a significant wage increase (of about 20%) in firms that trained in the previous 12 months, with large firms benefiting more than small firms. @hardy2022 find that assigning an apprentice to informal firms in Ghana increase firm size by about half a worker, and firm revenues by 5-15 percent per apprentice. While @crepon2019 looks at the impact of fee subsidies on firm's apprentice and employee stocks, they do not estimate the change in size or revenues that a firm can expect from hiring additional apprentices. Our study is thus the first, to our knowledge, to report the impact of dual system training on firm size and profits in SSA.

Finally, a number of studies have examined the financial arrangements between traditional apprentices and informal firms. @velenchik1995 studies the structure of apprenticeship contracts in small informal firms in Ghana, and identifies three main transactions between apprentice and firm --- apprentice wages, fees and allowances --- and distinguishes between two broad types of contracts, namely those with and those without training fees. She finds that firms that do not charge fees are smaller and tend to offer more specific training. @velenchik1995 and @frazer2006 also suggest that training fees may be a substantial source of financing for some firms, but do not provide estimates of the allowances, wages and other training costs that these fees are meant to offset. This study attempts to fill this gap.


<!-- @bolli2021  _participating companies are relatively large compared to the large share of micro and small companies in the Nepali economy, employ a relatively high share of high-skilled workers and face high labor market tightness....companies make average net benefits of about 17’000 NPR [137 USD - over 20 months]. These net benefits are substantially higher for the electrical engineering occupation (62’000 NPR) than for the mechanical engineering occupation (6’000 NPR), mainly due to the higher productivity of apprentices from the beginning onwards. The results suggest that apprentices create 270’000 NPR in productive value in the 20 months of workplace training. These benefits are reduced substantially by the costs that amount to 252’000 NPR. The wages of apprentices represent the most important cost component (56%), 9 Note that total benefits and instructor costs are calculated as the surfaces shown in Figure 6 and Figure 7. To derive total benefits and instructor costs based on values shown in Figure 8, it is important to remember that Figure 8 shows values after some time has elapsed rather than averages of a month. Hence, calculating total benefits and instructor costs based on these values consists of adding up the values but give a weight of 50% to the observations after three and 23 months. 25 followed by instructor costs (35%). Non-wage compensation is the least important cost component (9%). However, total benefits are larger than total costs, suggesting that the difference yields a net benefit for companies of 17’000 NPR._ -->

<!-- @deamesti2021 _by participating in dual training, the surveyed companies incur gross costs of $3,298,440 Chilean pesos (CLP) per year per student on average. This corresponds to US$4,650 or about 4.3 monthly wages of a skilled employee in the same area as the dual student. These gross costs consider four different types of costs: (1) training costs (57%), (2) student labour costs (28%), (3) administrative costs (13%), and (4) educational supplies (2%). Regarding the benefits, companies capture around $1,180,207 CLP per year per student. The benefits come from three sources: (1) the productive value that students contribute through their skilled and semi-skilled activities (79%), (2) savings in hiring (2%), and (3) savings in adjustment or induction (19%)._ -->

<!-- @chen2013 _have analysed the costs and benefits of apprenticeship training in a cooperation project between the Guangzhou Communications Technician Institute and Guangzhou Machine Tool Works Co, Ltd. using a calculation methodology also based on CBQ/QEK, but -- with regard to the cost benefit section -- adapted to the Chinese context. Data were collected within the framework of a three-year apprenticeship programme with 28 apprentices taking part. According to this field study, a company was able to obtain net benefits in each year of training, the first year being the most beneficial (for the exact figures, see Appendix, table A.5). This was due to the fact that leaners only received very little or sometimes even no training allowances in the first year of training. Moreover, it was discovered that the share of training allowances was about 90 per cent of the total cost of training, which easily explained the rather unusual result of high benefits in a first year of training in China._ -->


<!-- Dual "apprenticeship is a form of skill acquisition in the context of production, or on-the-job training. The apprentice, in addition to learning a trade, provides labor services to the firm. The relationship between the apprentice and the firm, therefore, includes two types of transactions, one being the provision of labor services by the apprentice to the firm, the other being the provision of training services by the firm to the apprentice. The two facets of the apprentice-firm relationship give rise to two components of the financial relationship between apprentices and firms. In reality, however, the monetary arrangements surrounding apprenticeship training in Ghanaian manufacturing are more complex. In addition to wage payments, many apprentices receive allowances for food, housing, clothing, or transportation. These payments are made both in cash and in kind but are identified by firms as distinct from wage payments. In larger firms, these allowances are paid to all workers and have been used as a way to generate nontaxable compensation.1 In smaller firms, allowances are often paid to apprentices even when wages are not. Historically, apprentices often lived with their masters during training. Although this practice has largely died out, provision of basic living expenses by the master continues to be a common feature of these contracts." [velenchikApprenticeshipContractsSmall1995]** -->

We find that, in general, all apprentices gain trade-specific human capital over the three observed years of training. However, we are unable to show that participation in dual system training contributes to additional learning. We find large variation in learning across trades, and the largest gains in human capital to be for youth with low learning scores at baseline. On the firm side, we find that the majority of MCs suffer net losses for each apprentice they train. The proportion of firms with negative net benefits from training, and the magnitude these losses, vary considerably depending on assumptions around allowances disbursed by the firms. Mean net benefits per apprentice range from a total `r mean(df$cb_II/605, na.rm = T)` \$US to `r mean(df$cb_V/605, na.rm = T)` \$US depending on the costs and benefits taken into consideration. Only 33%-45% of apprentices are estimated to generate a profit for their MC; a similar range captures the proportion of firms that are estimated to profit (generate a positive net benefit) from their training activities. Finally, we find that larger firms benefit the most, while smaller firms suffer the largest losses, suggesting that economies of scale contribute to the incentive structures surrounding firm training.

The paper proceeds as follows. Informal apprenticeship in Bénin, the CQP program, and the survey data used for the analysis are presented in Section \ref{data}. Results are presented in Section \ref{results}. Section \ref{conclusion} concludes.

\FloatBarrier 

# Data and Methods {#data}

## Country Context

Despite the relative stability of its democratic government and strategic importance as a transportation hub, Bénin (population approx. 12.1 million) performs poorly on many development indicators, ranking 158th out of 189 countries on the 2020 Human Development Index. Youth employment is a particularly pressing issue, with an alarming decrease in youth labor force participation in the past decade. As in other parts of SSA, secondary and tertiary school enrollment has seen a steady increase in the past two decades, with the predominant effect of displacing employment: according to the most recent labor force surveys, the youth employment-to-population ratio decreased by 22\%, from 40\% to 31\%, between 2011 and 2018, compared to an 8\% decrease for adults over the age of 25 over the same time period. Meanwhile, the share of youth neither in employment, education or training (NEET) increased from 17.2\% in 2011 to 35.1\% in 2018 (see Figure \@ref(fig:fig-enrolment)) --- one the highest rates in West Africa, and the world [@ilo2022].

```{r fig-enrolment, fig.cap="Rates of youth enrolment and inactivity: Bénin and SSA", fig.height=4, fig.width=8, , fig.pos='H'}
```

As enrollment in formal education has not translated to increasing rates of youth employment, interest in promoting alternative pathways to the labor force in Bénin has grown. Recent examples of investment in Bénin's apprenticeship system include $6.3 million from the World Bank's for the Benin Youth Employment Project (PEJ), completed in 2019, and a planned \$16.4 million dollar investment in strengthening the TVET system starting in 2020 [@worldbank2020].

In 2005, the government of Bénin announced a restructuring of traditional apprenticeship in the informal sector. Two national apprenticeship schemes were introduced: a formalization of the existing, firm-based apprenticeship program in the form of the _Certificat de Qualification aux Métiers_ (CQM) certification system, and the **dual system** _Certificat de Qualification Professionnelle_ (CQP) program, which combined in-firm training and classroom teaching. The three stated objectes were to (i) offer practical and theoretical training to youth under apprenticeship contracts in the craft sector (ii) train a high-performance labor force; and (iii) improve the productivity and profitability of workshops in the craft sector [@davodoun2011]. The government organs tasked with the administration of the CQP were the national TVET directorate (DETFP), the Direction of Test and Exam Services (DEC), in charge of the entrance and exit examinations for the CQP, and FODEFCA, responsible for procuring and distributing funding for the CQP [@nouatin2019]. The CQP began curriculum planning in 2005 with technical assistance from the French Development Agency (AFD) and the Swiss Agency for Development and Cooperation (SDC), among others, and became operational in 2008. In 2012, management of the program passed from Swisscontact, a Swiss NGO, entirely into the hands of FODEFCA [@nouatin2019].

Entry into formal technical and vocational education and training (TVET) begins after the completion of the second year of secondary school, or nine years of education. Yet across the country, the median number of years spent in the education system is four; only five percent of youth of secondary school age are enrolled in TVET [@ilo2021], in line with the six percent of young workers estimated to participate in formal TVET across SSA [@hofmann2022]. Thus, rather than formal TVET, it is informal apprenticeship that is the primary conduit into the labor market for early school leavers in Bénin, with as many as 300,000 young men and women estimated to be in training [@ilo2021].

Officially, the CQP recognizes "professional, technological and general skills acquired by the apprentice for the exercise of a professional activity" [@benin2010]. It also allows the pursuit of further technical and professional studies upon its completion. To participate in the CQP, applicants must (i) be at least 14 years old, unless otherwise authorized by the labor inspector; (ii) have a written apprenticeship contract that complies with labor laws; (iii) have completed at least 6 years of formal schooling; and (iv) pass a national entry examination [@kof2017]. Firm owners apply on behalf of the apprentices in their charge, generally through local craftsmen associations. 

The CQP is currently available for 13 out of the more than 300 trades listed in the craft sector: auto mechanics, motorcycle mechanics, air conditioning mechanics, tailors, masons, carpenters, metalworkers, electricians, and plumbers [@swisscontact2019]^[This selection of trades was based at least in part on existing trades from early experimental dual training programs to take advantage of existing training center infrastructure. The CQM is available for about 50 trades.]. Cost sharing for the CQP program is shared by the state and the apprentice, with the state financing body for dual apprenticeship, FODEFCA, officially taking on 90 percent of the training costs [@kof2017]. However, FODEFCA is largely reliant on external donor funding, and regular financing has been an issue for the program in recent years [@david-gnahoui2017].

<!-- With the goal of expanding the number of trades eligible for the CQP and CQM, the Béninese government, through the Youth Employment Project (PEJ) co-financed by the World Bank, has focused on expanding the programs to the crafts, tourism and food processing sectors in recent years.

*Dual apprenticeship* is a particular form of this institution, most popular in Germany and Switzerland but practiced across Europe, which combines on-the-job training with continued classroom-based general and trade-specific education. Generally, apprentices in the dual system spend between 20 and 30 percent of their training in the classroom, and train for two to four (most commonly three) years, depending on country and trade.  _Theory training enables young people to acquire general education as well as technical and vocational skills._ Due to the additional costs of classroom training, investments per apprentice will be higher, whether these are covered by the apprentice (in the form of a wage reduction), the firm, or state subsidies. Evidence from Germany suggests that dual training results in net costs in the time of training, which is recouped by firms when apprentices stay with the firm after training [@acemogluWhyFirmsTrain1998, @pfeifer2009]; though recent results suggest that net benefits during the training period are closer to neutral when the impact of hiring an apprentice on profits is compared to the hiring of similarly unskilled or semi-skilled workers on the labor market [@mohrenweiser2009].  -->

The financing of dual training comprises three main budget items: the firm/workshop, the training center and certification. While on-the-job training in the firm is paid for by the parents, training in the training center is largely financed through FODEFCA from various sources (national budget, donors, NGOs, etc.).

CQP qualification is attained upon the completion of a three to four year apprenticeship with a training firm in one of the designated trades or crafts and the passing of the annual national CQP examination. The final examination has a practical and a written component and is carried out by state representatives and local craftsmen. Upon successful completion, apprentices receive a nationally-recognized certificate. Certification upon successful completion of the CQP exam is allocated to the national budget via the Directorate of Examinations, DEC [@david-gnahoui2017].

Criticisms of the CQP include the lack of certified and accredited trainers at the training centers and unstable financing, which leads to high fluctuation in the number of applicants admitted to the program [@david-gnahoui2017; @ilo2020a].

<!---
_From an economic point of view, it seems obvious that a firm would hire the more apprentices the higher the net benefit of apprenticeship training amounts to. So far, most studies suggested that apprenticeship training results in net cost during the apprenticeship (e.g. Beicht et al. 2004). This would imply that companies need to be able to recoup these net cost by employing graduated apprentices as skilled workers and paying them a wage lower than the value of their marginal productivity, because otherwise they would not o↵er an apprenticeship programme. Themostrecent study of the Swiss apprenticeship system shows, unlike other studies, that about two thirds of the training firms find it profitable to train apprentices (Schweri et al. 2003)._
--->

## Sampling

The data for this study was collected in two separate surveys. The first consisted of interviews with apprentices who had applied to the 2019 cohort of the CQP program; the second was conducted with the owners of their respective training firms. To reduce travel distances and allow for subsample analyses across trades, apprentices were randomly selected from a subsample of apprentices: those who had applied to train in electrical installation, carpentry, masonry, metalwork (primarily welding of gates for living compounds), or plumbing (out of a total of 13 participating trades) and who were training in the south of Bénin.

In addition to questions regarding training practices and firm performance, master craftsmen were interviewed on the subject of specific apprentices training at their firm. These apprentices were either participants in the CQP program, unsuccessful applicants to the program, or non-applicants. As the apprentice survey consisted only of applicants to the program, the only data on non-applicants comes from master craftsmen. To limit the duration of the firm interviews, apprentice-specific questions were only posed for a randomly chosen subset of apprentices for firm owners with high numbers of trainees. This sampling procedure is summarized in Table \@ref(tab:sampling) below.

| CQP Status | Explanation | Apprentice Survey Sampling | Firm Survey Sampling |
|---|---|---|---|
| Selected | Already in apprenticeship, applied via master craftsmen. Passed exam and selected. | Random sampling from list of all CQP  applicants in five chosen trades and geographic region (Southern Bénin) | If more than one CQP applicant training in firm, at most two apprentices were  randomly selected at baseline. |
|   |
| Not selected | Already in apprenticeship, applied via master craftsmen. Not selected due to exam score or lack of training center. | Random sampling from list of all CQP  applicants in five chosen trades and geographic region (Southern Bénin) | If more than one CQP applicant training in firm, at most two apprentices were  randomly selected at baseline. |
|   |
| Did not apply | Did not apply to CQP. Trained as traditional apprentice. |  N/A | Firm owner listed *up to* 5 apprentices who did not apply to CQP. One randomly selected at baseline.|

Table: (\#tab:sampling) Apprentice Sampling

The baseline wave for the two surveys was collected in July-August 2019. The apprentice survey included questions on training characteristics, employment outcomes, skill and competence assessments and apprentice assessments of training quality, while the firm survey included questions on workforce composition in the firm, wages, costs and revenues. In addition, we surveyed all MCs about the firm's training practices and expenditures for training, as well as their perception of individual apprentices’ skills, experience, diligence, efficiency, learning ability, and so on. Data on `r (df %>% group_by(wave) %>% summarise(n_distinct(IDYouth)))[1,2]` apprentices working for `r (df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[1,2]` unique firms was collected at baseline. Descriptive statistics for apprentices and firms is shown in Table \@ref(tab:tbl-desc) below. 

```{r, tbl-desc, echo=FALSE}
```

Summary statistics from the baseline survey show the sample to be predominantly male youth who, though of average age for an apprenticeship at `r mean(baseline$age, na.rm = T)` years [@ilo2022], are significantly more educated than is typical for traditional apprentices, with over half having completed at least some secondary schooling. Apprentices selected for participation in the CQP are more educated than those not selected. Though all applicants applied to the same cohort of the CQP program, they came in having more experience than the required six months: the average apprentice had `r mean((df %>% filter(SELECTED == 3))$baseline_duration, na.rm = T)` years of experience at the time of application. Non-applicants have the lowest educational attainment and are the least experienced at baseline, suggesting that master craftsmen send most able apprentices to stand for the entrance exam. Finally, it is notable that CQP applicants have over three years of apprenticeship experience on average, though the CQP entrance requirement only stipulates six months. This again suggests a selection mechanism for more experienced/able applicants that may indicate that the CQP is considered more of a continuing education program among craftsmen than an entry-level apprenticeship program.

The majority of training firms are small workshops comprising the firm owner --- the master trainer --- and several apprentices. Two firm sizes are shown in Table \@ref(tab:tbl-desc): those stated directly by the firm owner in response to the question, "How many people (including you and your apprentices) are currently working in your business?" and those calculated by the author by summing the number of apprentices, partners, paid and unpaid family workers, and occasional workers engaged by the firm owner. Using self-reported size, `r sum(prop.table(table(baseline$firm_size_bins_reported))[1:2])*100`% of firms employed a total of five or less (including the owner) at baseline and `r sum(prop.table(table(baseline$firm_size_bins_reported))[1:3])*100`% of firms employed no more than ten. Using the calculated firm size, `r sum(prop.table(table(baseline$firm_size_bins))[1:2])*100`% of firms employed a total of five or less (including the owner) at baseline and `r sum(prop.table(table(baseline$firm_size_bins))[1:3])*100`% of firms employed no more than ten. Thus, training firms in the sample are small, in line with observations from the informal sector in Ghana [@frazer2006, @velenchik1995]. Moreover, the workforce of the average firm in the sample is dominated by apprentices: at baseline, the average firm employs about four apprentices for every other type of employee (`r mean((firms %>% filter(wave == 0, firm_size > 1))$FS6.1/((firms %>% filter(wave == 0, firm_size > 1))$firm_size-1), na.rm = T)*100`% of calculated firm size at baseline and `r mean((firms %>% filter(wave == 1, firm_size > 1))$FS6.1/((firms %>% filter(wave == 1, firm_size > 1))$firm_size-1), na.rm = T)*100`% at endline).

The endline survey was conducted in August-September 2021. It is quite common for studies of training programs to be affected by high rates of attrition [@mckenzie2017]. Overall apprentice attrition in our sample, at `r round((1-(df %>% group_by(wave) %>% summarise(n_distinct(IDYouth)))[2,2]/(df %>% group_by(wave) %>% summarise(n_distinct(IDYouth)))[1,2])*100,1)`%, is clearly very high, even when compared to studies in similar contexts: for instance, @crepon2019 and @hardy2019 both report youth attrition of around 10%. It is driven both by youth non-response and firm non-response^[Data on apprentices who had not applied to the CQP was only obtained from firm owners, and thus only subject to firm attrition.]. However, dropout rates suggest that survey attrition is not caused predominantly by apprenticeship leavers: even when considering as dropouts apprentices whose reason for discontinuing training is unknown, just `r sum(endline$grad == "Dropped out" | endline$grad == "Unknown", na.rm = T)/nrow(endline)*100`% of all apprentices for whom data was collected at endline left their program before graduation. `r sum(endline$grad == "Graduated", na.rm = T)/nrow(endline)*100`% had graduated after three years, and the majority, `r sum(endline$grad == "Still training", na.rm = T)/nrow(endline)*100`%, were still training. Graduation rates are highest for electrician apprentices, at `r sum((endline %>% filter(FS1.11 == 5))$grad == "Graduated", na.rm = T)/nrow((endline %>% filter(FS1.11 == 5)))*100`%, but comparable across trades. However, graduation rates were much lower for participants in the CQP than non-participants, potentially caused by Covid-19 related training delays. As we were observing the cohort that began training in 2019, it is consistent with an apprenticeship duration of 4 years that the majority of the sample was still training after three years. Program dropout is quite low compared to similar studies, e.g. @crepon2019 who report 31.2% dropout for dual apprenticeships and 32.5% for traditional apprenticeships. Table \@ref(tab:tbl-attritionapps) in the Appendix gives no indication of attrition bias, as there is no systematic difference in the sample composition by CQP participation status, trade, or relevant socioeconomic variables.  

Similarly, of `r (df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[1,2]` firms interviewed at baseline, only `r (df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[2,2]` could be contacted at endline, for an overall firm attrition rate of `r round((1-(df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[2,2]/(df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[1,2])*100, 1)`%. Table \@ref(tab:tbl-attritionfirms) in the Appendix likewise suggests that firm attrition was not correlated with key firm characteristics such as trade or the number of CQP applicants (selected or non-selected).

Though dual training is predicated on classroom teaching about once a week, we found that external training (classes or training that took place outside of their master's workshop) was not limited to CQP participants. At endline, `r mean(as.numeric(endline$YS4.43), na.rm = T)*100`% of apprentices reported participating in such training in the preceding three months. However, only `r mean(as.numeric((endline %>% filter(SELECTED == 1))$YS4.43), na.rm = T)*100`% of CQP participants reported doing so, despite external training being a constituent and necessary component of the CQP program's dual training structure.

The majority  (`r sum(df$YS4.48[!is.na(df$YS4.48)]<4)/length(df$YS4.48[!is.na(df$YS4.48)])*100`%) of apprentices reported that the training center they attended was within one hour of traveling distance. Among apprentices who attended classes or training outside their master's workshop, nearly three quarters <!--`r 64/(64+23)*100`%--> reported that the training took place in a training center (the rest said it took place in another workshop).

By combining time trained at the time of the survey with the expected graduation date, we find that apprentices expect their training to last `r mean(df$projected_duration, na.rm = T)` years, with a standard deviation of `r sd(df$projected_duration, na.rm = T)` years, in line with our assumption of four-year apprenticeships. Metalworking apprenticeships reportedly last almost a year longer than the other trades in the sample.

Apprentices also explained their motivation for starting an apprenticeship. The most frequently cited motive was interest in the trade, followed by the promise of good earning opportunities and the insistence of the parents. They were also asked why they had chosen the craftsmen they were training with; the majority replied that they were attracted by the patron's reputation, or that their parents had made the choice in their stead. As found in previous studies of informal apprenticeship in SSA, most apprentices express a wish to start their own business after graduation (`r sum(df$YS4.53[!is.na(df$YS4.53)]==2)/length(df$YS4.53[!is.na(df$YS4.53)])*100`%) followed by seeking employment with a different workshop (`r sum(df$YS4.53[!is.na(df$YS4.53)]==1)/length(df$YS4.53[!is.na(df$YS4.53)])*100`%). 

<!-- Apprentices report working `r round(mean(df$a_weekly_hours, na.rm = T),2)` hours in the previous week, compared to `r round(df %>% group_by(FS1.2) %>% summarise(x = mean(firm_weekly_hours)) %>% select(x) %>% colMeans(na.rm = T),2)` hours reported by the patron. -->

## Estimating Apprentice Benefits {#appmethod}

We first examine the benefits accruing to apprentices over the observed time period of three years. These benefits can be separated into human capital gains and material benefits.

Human capital gains are measured using a set of trade-specific scores measured separately for each apprentice at baseline and endline. These amount to a simplified version of the "task approach" utilized in the technological change literature [see @dicarlo2016; @crepon2019]. Unlike the general tasks used to measure skill  in the task approach, however, we measure craft knowledge using a short test designed using CQP curricula, and experience using a short roster of tasks drafted in collaboration with local craft experts and practitioners, similar to @hardy2019. Firms were asked to assess apprentices on this series of trade-relevant tasks, while apprentices were given the short knowledge test as a small part of their interview at baseline and endline. Similar to the task approach, this method allows for worker-level measurement of ability and experience based on tasks performed, as each apprentice receives a score in each of the three dimensions. 

Competence and experience scores are the percentage of tasks in which apprentices are deemed competent or experienced. They are based on a set of 10 to 15 tasks for each trade in question that apprentices should master in the course of an apprenticeship in their selected trade that were selected with the help of active local craftsmen (shown in Appendix \@ref(hcqs)). Apprentices are evaluated on each task by their master trainer^[Apprentices were asked to self-evaluate their competence at endline using the same metric. Self-evaluation was not initially planned and thus unavailable at baseline] on a binary scale: they are either competent at realizing a given task (competence metric) or have already realized a given task in the past (experience metric), or not.

The knowledge score is based on a short battery of questions drawn from the official competence charts for each trade and posed directly to the apprentices. Each question was a multiple choice question, and between 4 and 5 knowledge questions were posed to each apprentice. The knowledge questions are reproduced in Appendix \@ref(knowqs). Because apprentices who did not apply to the CQP were not interviewed directly, the knowledge score was only measured for CQP applicants.

We also use regression analysis to examine the impact of dual training on our various measures of apprentice learning outcomes. We use the specification

$$ y_{it} = a+\sum_{j}\text{status}_{ij}+CQP_{i}\text*{wave}_t+\mathbf{X}_{it}+\mathbf{Z}_{jt}+u_{it} $$ 
where $y_{it}$ is the outcome for apprentice $i$ at time $t$, $\text{status}_{ij}$ corresponds to apprentice status $j$ of apprentice $i$ in the context of the CQP program for: either successful applicant, unsuccessful applicant, or non-applicant. $\mathbf{X}_{it}$ is a column vector of apprentice characteristics, $\mathbf{Z}_{jt}$ is a column vector of training-related training characteristics, $a$ is a constant, and $u_{it}$ is an error term. $CQP_{i}\text*{wave}_t$ is a dummy for CQP apprentices interacted with survey wave, which identifies any gains in learning outcomes that can be linked with participation in the CQP.

Material benefits accruing to apprentices amount to total fees paid less allowances received. Fees are typically paid by the parents directly to the MC at the beginning and end of an apprenticeship. Allowances are transfers from the MC to the apprentices and cover general expenses, in which case they are often referred to as "pocket money" or "chop money", as well as specific expenses such as meals or transportation costs. These are described in greater detail in the next section.

## Estimating Firm Benefits {#firmmethod}

\input{tables/tbl-cbmodels}

Firm benefits were calculated using the accounting approach [@gambinlynn2013; @muhlemann2014], which relies on survey data from training firms to identify and quantify the costs and benefits arising from training provision. It has only recently started being applied in lower-middle income countries [@renold2018; @bolli2020; @bolli2021]. The accounting approach subtracts the costs of training from its benefits and is a simple way of estimating the value that training apprentices generates for the training firm. In a second step, we use regression analysis to assess the impact of hiring additional apprentices, both traditional and CQP dual trainees, on firm outcomes.

The components of the accounting approach used in this study are summarized in Table \@ref(tab:tbl-cbmodels). Firms receive two primary benefits from training apprentices: apprenticeship fees and the apprentices' productive contributions to the firm. Training fees can be paid in full before the commencement of training or split into payment at the beginning, during, and at the conclusion of training [@velenchik1995]. Five categories of fees were reported by both apprentices and MCs: entry fees, formation (or general training) fees, liberation (or graduation) fees paid at the conclusion of training, fees as compensation for the materials and equipment used in training, contract fees, and application fees\footnote{Fees are often paid in kind rather than in cash.}. Fees were reported as the total paid for the entirety of the apprenticeship; we assume four-year apprenticeships to estimate annual ammounts.

The second benefit of training for firms, apprentices' net productive contributions, were not reported explicitly by the firm owners and thus needed to be estimated with the aid of several assumptions. First, we assume the competitive model of labor markets (with heterogeneous wages), in which workers are paid the marginal product of their labor. Second, we assume apprentice productivity is equal to that of an untrained employee with no more than a primary education for the first two years of training, and increases to that of trained employee for the final two years\footnote{This is a simplification of the approach used by @bolli2020, in which apprentice productivity is estimated to increase linearly from that of an unskilled worker to that of a skilled worker between defined points in their training. A popular alternative to this approach involves eliciting specific tasks performed by apprentices and estimating costs savings based on the wages paid to workers who would otherwise be responsible for said tasks [@hauschildt2018]. Our firm-apprentice data did not cover specific tasks and is thus not equipped to carry out such an analysis.}, and we use detailed wage information reported by firm owners to estimate the average annual productivity over the course of a four-year apprenticeship. Under these assumptions, the annual productive value generated by apprentice work amounts to the average of these two wages.

Costs of apprenticeship for the firm are categorized into three categories: allowances, training expenses, and lost trainer productivity.

Allowances are disbursed irregularly by the firm owner for small expenses such as travel and meals. These are reported by firms at the apprentice level (separate reported allowances for each apprentice). To estimate total annual allowance expenditures per apprentice, we thus sum over all allowance categories and assume that apprentices work 20 days per month; the extrapolated monthly sum is then multiplied by the number of months the training firm was operating in the past year to arrive at an annual estimate for each apprentice.

To identify the costs accruing to the firm, owners were asked to identify any costs directly or indirectly related to their training activities. These costs can be divided into two categories: Equipment costs comprise all costs for physical infrastructure necessary for training: raw materials such as cement, lumber, or scrap metal used in the course of training; training equipment such as workbenches, toolkits, or other machines purchased or rented specifically for training purposes, rent for training facilities if training was not conducted exclusively in the firm owner's workshop, and books and any other training materials. 

Lost trainer productivity is, like apprentice productivity, estimated using wage data. Hourly wages for skilled employees were calculated from monthly wage data and multiplied by the number of hours that the workshop stopped all productive activities to train apprentices in the previous week, as reported by MC. This estimate is burdened by the largest number of assumptions: it is uncertain whether all employees who train apprentices in the firm (a number reported by the MC) stop work entirely while the workshop takes a break to train; whether the majority of lost productivity occurs during these breaks, or in the otherwise normal operation of the firm during which they must also tend to the apprentices. Moreover, the total duration of these breaks in the past week is a very small sample from which to extrapolate to annual costs. Lacking a better method, we report these estimates as the final cost component.

<!-- Firms estimated the hours trained on the last day the firm stopped all productive activities to train apprentices, as well as the number of days per week that such training occurred. We use this information as the basis for our estimation. We extrapolate weekly hours of training (hours trained on previous day of training x days trained per week) to annual hours by assuming four work weeks per month and multiplying by the number of months the firm owner reported being open in the previous year. As when estimating apprentice productivity, we set reported monthly wages for skilled workers (wage employees who had trained with the current firm) equal to trainer productivity and divide by the approximate hours worked in the past month to arrive at approximate hourly wage per trainer (assuming four work weeks per month and using firm-reported days open last week and hours worked on the last day). Finally, we multiply by the number of trainers and divide by the number apprentices per firm to arrive at an estimated cost, per apprentice, in terms of total foregone employee productivity resulting from training activities. -->

Net benefits were calculated by subtracting some or all of the costs listed above from some or all of the benefits. The simplest model, Model I in \@ref(tab:tbl-cbmodels) above, reduces the value of training accruing to the firm to the direct material incomes and expenditures related to training; namely, apprenticeship fees from the apprentices and their parents, an approximation of training costs, and the total reported allowances disbursed to the apprentice. Models II and III add training expenses and estimated apprentice productivity, respectively, while Model IV combines all four components. Model V includes all costs and benefits, including lost trainer productivity.

<!-- $$ \textbf{Model I:} $$ -->
<!-- $$\textit{annual net benefits}_i = \frac{\textit{fees paid}_i}{4} - \textit{apprentice allowances}_i - \frac{\textit{total training costs}_j}{N_j}$$ 

\noindent for each apprentice $i$ in firm $j$, where $N_j$ is the total number of apprentices in firm $j$.-->

<!-- $$ \textbf{Model V:} $$ -->

<!-- $$ \begin{aligned} \textit{annual net benefits}_i =  &\frac{\textit{fees paid}_i}{4} - \textit{apprentice allowances}_i - \frac{\textit{total training costs}_j}{N_j} + \\ & \textit{estimated apprentice productivity}_i - \\[1em] & \frac{\textit{estimated foregone productivity per trainer}_j}{N_j} \end{aligned} $$

\vspace{.5cm}
\noindent for each apprentice $i$ in firm $j$, where $N_j$ is the total number of apprentices in firm $j$.  -->

Finally, we estimate the effect dual training on firm size and profits by running a pooled OLS regression taking the form

$$ y_{it} =  a+CQP_i+apprentices_{it}+\mathbf{X}_{it}+u_{it}, $$

where $y_{it}$ is the outcome of interest, $\mathbf{X}_{it}$ is a matrix of additional covariates for firm $i$ in wave $t$, and $u_{it}$ is an error term. $CQP_i$ is the number of CQP applicants who were accepted into the 2019 cohort of the program, while $apprentices_{it}$ controls for the total number of apprentices training with the firm (and in contrast to $CQP_i$ is a time-varying measure).

\FloatBarrier

# Results {#results}

##	Impact of Informal and Dual Training on Individuals {#appbenefits}

```{r changes, include = F}
changes <- df  %>% select(IDYouth, comp_all_trades, exp_all_trades, skills_all_trades, wave) %>% pivot_wider(names_from = wave, values_from = c(comp_all_trades, exp_all_trades, skills_all_trades)) %>% mutate(comp_diff = comp_all_trades_1-comp_all_trades_0, exp_diff = exp_all_trades_1-exp_all_trades_0, skills_diff = skills_all_trades_1-skills_all_trades_0)
```

```{r tbl-skills}
```

First, we investigate whether dual training was successful in realizing its primary objective --- increasing the human capital of the apprentices. To do so, we study the changes in the three human capital indices described in Section \@ref(appmethod) over the observed training period of three years. The changes in the human capital index scores presented in Table \@ref(tab:tbl-skills) indicate that informal apprenticeship training is successful in improving sector-specific human capital of the youth in our sample, both for dual training participants and traditional apprentices: overall competence scores increased by `r mean(changes$comp_diff, na.rm = T)/sd(baseline$comp_all_trades, na.rm = T)` baseline standard deviations (SDs), experience scores by `r mean(changes$exp_diff, na.rm = T)/sd(baseline$exp_all_trades, na.rm = T)` SDs, and knowledge by `r mean(changes$skills_diff, na.rm = T)/sd(baseline$skills_all_trades, na.rm = T)` SDs. Significant improvements in competence and experience are observed for apprentices who participated in, unsuccessfully applied to, and did not apply to the CQP alike, though apprentices who did not apply to the CQP show the largest gains in competence and experience as assessed by their MC. This result is in line with the observation that MCs appear to send relatively experienced youth to apply for dual training. 

A paired t-test also indicates significant improvements in competence and experience between baseline and endline across all trades (improvement in competence for plumbing apprentices is marginally insignificant at standard significance levels). On the other hand, gains in the knowledge metric are not statistically significant for any trades except plumbing - hence, the overall increase in knowledge is driven by results from a single trade. As the average knowledge scores at baseline were significantly lower for plumbing apprentices than for apprentices in other trades, this result may indicate a shortcoming in the metric itself, which was composed of only up to five questions which did not seem to pose a major challenge for most apprentices not in the plumbing trade. Improvement in knowledge was marginally significant (p < 0.10) for CQP participants but not for unsuccessful CQP applicants (p < 0.5). 

Although mean human capital accumulation is higher for participating CQP apprentices than non-selected CQP applicants across the three indices, this does not translate into statistically significant differences between the two group (Table \@ref(tab:tbl-skillschangebycqp) in the Appendix). Nor do we observe a significant effect of dual training on the competence and experience indices when apprentices who did not apply to the CQP are added to the control group. To check if apprentices agree with their MC's assessments of their ability, a self-assessment was included in the apprentice survey at endline. Table \@ref(tab:tbl-compexp2) in the Appendix suggests that the two assessments are in general agreement for both the experience and competence index.

<!-- The three metrics used to measure apprentice progress are not comprehensive, in that they do not measure outcomes such as the physical and psychological well-being of apprentices, "soft skills", etc. Thus, we also examine subjective measures of apprenticeship quality captured by a 5-point Likert scale rating of the following characteristics of the firm, as perceived by the apprentice: physical safety of the workshop, the master trainer's knowledge, the master trainer's treatment of apprentices, the treatment of the apprentice by other employees and apprentices, apprenticeship benefits, apprenticeship working hours, the quality of training, and the quality of the equipment and machinery in the workshop. We observe generally positive levels of satisfaction with training quality, ranging from an average of 4 (somewhat good) to 5 (very good) on all categories except physical safety at baseline (average = 3.96) and working hours at endline (3.66). Subjective ratings are also generally constant over time. One exception is satisfaction with working hours, which decreases from an average rating of 4.07 to 3.66 - indeed, we observe an increase in reported working hours from 42 to 45 average hours worked in the past week for apprentices who applied but did not participate in the program (while CQP participants work hours remained stable). Generally, we observe no significant differences in satisfaction between CQP participants and apprentices who applied to the program but did not participate. -->

<!-- To summarize, satisfaction is generally high and stable among all apprentice throughout their training period, though apprentices not participating in the CQP report a sharp increase in working hours towards the tail end of their training. -->


``` {r, tbl-appreg, results = 'asis', fig.align = 'center'}
```

\noindent The estimated effects of training presented in Table \@ref(tab:tbl-appreg) confirm that apprentice human capital increases after three years, as measured by the three indices, and that apprentices who do not apply to the CQP program score lower to begin with. Higher baseline experience and competence scores for CQP applicants suggest that trainers send their more able apprentices to apply for dual training. The selection process for the program itself, however, does not favor more experienced apprentices according to our metrics (i.e. is as good as random).

Participation in the CQP program is shown to have no detectable effect on human capital accumulation; in fact, when controlling for individual effects, we find that apprentices who did not apply to the CQP make larger gains on the competence and experience indices relative to non-selected CQP applicants, while CQP participants do not (columns (3) and (6) in Table \@ref(tab:tbl-appreg)). The regression estimates also suggest that apprentice experience and the number of other apprentices training in the firm are correlated with somewhat higher index scores. Table \@ref(tab:tbl-appreg2) in the Appendix shows the same regression estimates using a restricted sample of CQP program applicants (successful and unsucessful) only, with the same result: there is no detectable effect of dual training on apprentice experience or competence. 

```{r, tbl-netappbenefits, echo=FALSE}
```

\noindent We also estimate the apprentices' net cost of training, taking into account material costs in the form of the various training fees, and material benefits in the form of various allowances disbursed by the MC over the course of the training period, as outlined in Section \@ref(appmethod). Table \@ref(tab:tbl-netappbenefits) showing annual fees and allowances reported by both the apprentice and MC indicate that trainees in fact receive more in allowances from their trainer than they (or their parents) pay in total fees (assuming a four-year training duration). Formation, or general training, fees comprise the largest sum transfer from the apprentice to the MC. Apprentices report significantly higher fees than the MC, for the training fee in particular. Firms may underreport fees to avoid accusations of gauging, but are at the same time likely to have more direct knowledge of all fees than apprentices, whose parents and relatives usually pay the craftsmen directly. 

Overall, apprentices are estimated to receive about 80,000 FCFA (134 \$US) more in allowances than they pay in fees annually, according to their own estimates; MCs estimate this number to be as much as 210,000 FCFA annually (345 \$US). MCs report that about half of all allowances are in the form of general "pocket money".

CQP participants do not report higher fees or allowances than unsuccessful applicants, while MCs report significantly different sums paid by non-applicants in fees, especially for formation (general training) fees. We observe no statistically significant difference between allowances disbursed to CQP participants relative to the other two groups.

<!---
Most classroom training takes place in training centers that are less than an hour away from the apprentices' domicile. 

According to survey data, about a quarter of respondents say training does not take place at a training center, but rather at another firm. At endline, when the answer choice was first included, 50% of apprentices reported attending a technical high school and 44 reported attending a training center. 70% report continuing CQP classes during the Coronavirus pandemic.
--> 

## Impact of Informal and Dual Training on Firms  {#firmimpact}

```{r, tbl-netbenefits, echo=FALSE}
```

\noindent Next, we study the net benefits accruing to firms from apprenticeship training, and whether the presence of dual-system apprentices in the firm impacts productivity relative to firms with fewer or no CQP participants. To measure benefits in the accounting sense, we refer to two categories of benefits --- fees and apprentice productivity --- and three categories of costs --- allowances, training costs, and lost trainer productivity --- as measured at baseline and described in detail in Section \@ref(firmmethod). Because productivity and training costs are particularly sensitive to assumptions and prone to non-response issues, these categories are combined in different permutations to give the five "models" referenced in Table \@ref(tab:tbl-cbmodels).

We first analyze net benefits per apprentice. Table \@ref(tab:tbl-netbenefits) estimates the annual costs and benefits per apprentice for each category and subcategory for all sample apprentices in the first column, using baseline responses; for successful and unsuccessful applicants to the CQP in the second and third columns, respectively; and for non-applicants in the fourth column. It shows that apprentice productivity (a benefit to the firm) and apprentice allowances (a cost to the firm) are the most important factors in determining whether an apprentice contributes a net benefit to the training firm. Estimated net benefits per apprentice range from `r round(mean(baseline$cb_I/605, na.rm = T),0)` \$US to `r round(mean(baseline$cb_V/605, na.rm = T),0)` \$US per apprentice per year, depending on the model used. The sign of the estimate depends on the inclusion of apprenticeship productivity (not included in Models I and II; included in Models III-V). Training fees are significantly higher and allowances are significantly lower for apprentices who did not apply to the CQP, resulting in higher net benefits for this subgroup.

Next, we look at the individual categories in more detail. Formation fees represent the largest single fee paid to the firms and account for over half of total fees paid. Other minor fees cover the provision of equipment and materials, application fees (pertinent for the CQP, as the master trainer must submit paperwork in their apprentices' stead), and initiation and graduation fees. In total, apprentices report paying about 175,000 FCFA (280 \$US) in fees for training, while firm owner report around 160,000 FCFA (255 \$US) in fees per apprentice.  This indicates a minor increase in the costs of training in Bénin over the past two decades: @walther2007 reports total fees ranging from 50,000 to 150,000 FCFA (96-290 \$US, inflation adjusted). Though generally unregulated, in some cases professional associations and public authorities step in to regulate fees, particularly those levied for initiation and graduation ceremonies. In contrast to the dual training program studies by @crepon2019, the CQP program did not eliminate fees or directly subsidize apprentices.  

Fees: We observe significant differences between apprentice types for overall fees (driven by higher training fees for dna) and contract fees (higher fees for CQP applicants), but not total benefits. A t-test for differences between the CQP participants and CQP applicants shows that formation fees are significantly higher (but only \$9 difference annually) for unsuccessful CQP applicants; otherwise benefits are not measureably different (though nearly \$400 higher for CQP non-selected, driven by productivity.)

Apprentice productivity: trainee productivity is calculated using firm-level wage information, an approach similar to apprenticeship cost-benefit studies by @wolter2015, @muhlemann2018, @bolli2020, and @bolli2021 and described in Section \@ref(appmethod). The annual productivity estimates reported in Table \@ref(tab:tbl-netbenefits) are based on the number of months the training firm reported being open in the previous year and are an order of magnitude higher than the estimated annual benefits from training fees. Because many small firms did not report the wage data necessary to estimate apprentice productivity, however, apprentice productivity can only be estimated for about a quarter of the sample. The estimated productivity for CQP apprentices is lower than for other trainees, but the difference is not statistically significant.

Allowances: annual allowances are estimated by multiplying the total firm-reported daily allowances for each apprentice by an assumed 20 days worked a week and the reported number of months open in the previous year. Alternative estimates using different workload assumptions and apprentice responses are shown in Table \@ref(tab:tbl-allowancebounds) and Table \@ref(tab:tbl-allowboundsapp).

Table \@ref(tab:tbl-netbenefits) shows that allowances disbursed to CQP participants and CQP applicants are estimated to be significantly higher than those disbursed to non-applicants. This suggests that older and more experienced apprentices receive higher remuneration (i.e. are already more productive when they apply to the program). (There is no difference between CQP apprentices and unsuccessful CQP applicants). CQP apprentices train, on average, in firms which pay less for equipment (whether only unsuccessful CQP applicants are all other apprentices are the control group), suggesting that the dual system may be preferred by firms with less access to machinery and capital.

Training costs: Table \@ref(tab:tbl-netbenefits) estimates that firm training expenditures, including materials and equipment costs, are an order of magnitude smaller than total allowances on a per-apprentice basis. There is no difference between types of apprentices. Firms report training costs for each category the past month; to estimate annual training costs per apprentice per year, the reported costs are thus divided by the number of apprentices training in the firm and multiplied by the number of months the firm was open in the previous year. Annual sums for training costs are reported at the firm level and thus normalized by the number of apprentices training in the firm, yielding approximate costs per apprentice per year.

The various costs of training reported by the MCs are depicted per apprentice in Figure \@ref(fig:fig-costspie) in the Appendix. All costs total to approximately 240,000 FCFA, or 400 \$US, per apprentice per year. The largest expenses are associated with allowances and raw materials (which are particularly high in the carpentry trade). They also in line with the upper range of costs from @david-gnahoui2017, who, citing Zinsou, 2012, reported total costs of 100,000 to 250,000 FCFA (\$165-\$413) for a complete CQP training program in 2012. Figure \@ref(fig:costspie2) in the Appendix accounts for the number of apprentices training in firms and reports mean total training expenditures per firm. These total, on average, 1.4 million FCFA, or 2,325 \$US, per firm, corresponding to over half of the reported annual firm revenues (`r round(mean(df$FS4.7/605*df$FS4.1, na.rm = T), 0)` \$US) and about 135\% of reported total annual costs (`r round(mean(df$FS5.1/605*df$FS4.1, na.rm = T), 0)` \$US), as averaged over the two survey waves.

Trainer productivity: similar to apprentice productivity, lost trainer productivity due to work stoppages is estimated using firm-level wage information and the number of months the training firm reported being open in the previous year. It also extrapolates from the reported number of hours of training on the last day the firm stopped all activities to train apprentices and the estimated number of days per week such a stoppage takes place (multiplied by an assumed four weeks per month that the firm is open).

Thus estimated, lost trainer productivity per apprentice is on the same order as the total other training costs (about 30 \$US per year) and an order of magnitude smaller than the allowances disbursed by the MCs to each apprentice. Lost productivity is comparable for CQP participants, applicants and non-applicants.

Net benefits: estimated mean benefits per apprentice ranges from `r mean(df$cb_I/605, na.rm = T)` \$US per year to `r mean(df$cb_III/605, na.rm = T)` \$US per year, depending on the model used. Whether our estimate returns positive or negative net benefits hinges on the inclusion of apprentices' productive contributions: Models I and II exclude this factor, resulting in drastically lower estimated benefits. 

The distributions of per-apprentice net benefits have long left tails for Models I and II and long right tails for Models III-V (plotted with their means in Figure \@ref(fig:fig-apphist) in the Appendix). The left tails are a consequence of the high number of apprentices in a number of firms generating unrealistic annual allowance totals; these are more than compensated by apprentice productivity estimates when included in Models III-V, skewing the distribution to the right for these models. 

For models excluding apprentices' productive contributions to the firm (Model I and Model II), we find that apprentices who applied to the CQP program (Selected and Not Selected) are significantly more costly to train (incur higher net costs) on average than non-applicants, on account of the higher allowances MCs report they receive. There is no difference between Selected and Not Selected.

Estimates for Models III-V are only available for a small number of apprentices due to the requirement that net benefits only be computed when data for all cost and benefit categories included in the respective model are available. The variance in benefits is high, ranging `r max(baseline$cb_III/605, na.rm = T)` \$US to net costs of  `r min(baseline$cb_V/605, na.rm = T)*-1` \$US per apprentice per year. Mean net benefits per apprentice per year range from `r mean((baseline %>% filter(SELECTED == 1))$cb_V/605, na.rm = T)` \$US to `r mean((baseline %>% filter(SELECTED == 1))$cb_III/605, na.rm = T)` \$US for CQP apprentices, `r mean((baseline %>% filter(SELECTED == 0))$cb_V/605, na.rm = T)` \$US to `r mean((baseline %>% filter(SELECTED == 0))$cb_III/605, na.rm = T)` \$US for unsuccessful CQP applicants, and `r mean((baseline %>% filter(SELECTED == 3))$cb_V/605, na.rm = T)` \$US to `r mean((baseline %>% filter(SELECTED == 3))$cb_III/605, na.rm = T)` \$US for non-applicants. Though large in magnitude, the differences in means between the three groups are not statistically significant for Models III and IV due to the limited sample and high variance. Model V, which includes all cost and benefit categories, generates statistically different estimates for net benefits for the three categories: CQP apprentices are nearly break-even, while unsuccessful CQP applicants generate benefits of `r mean((baseline %>% filter(SELECTED == 1))$cb_V/605, na.rm = T)` \$US and non-applicants generate `r mean((baseline %>% filter(SELECTED == 3))$cb_V/605, na.rm = T)` \$US. This is only based on a total sample of 28 observations, however.

CQP participants generate significantly lower benefits when unsuccessful CQP applicants only are used as the comparison group according to Models IV and V (\@ref(tab:tbl-appnetbenefitsnodna).

Only `r round(sum(baseline$cb_II>0, na.rm = T)/sum(!is.na(baseline$cb_II))*100, 0)`-`r round(sum(baseline$cb_I>0, na.rm = T)/sum(!is.na(baseline$cb_I))*100, 0)`\% of apprentices generate positive net benefits when productivity is left unaccounted for; this rises to about 70\% (at baseline) for Models III-V. According to Model I, `r round(sum(baseline$cb_I>0, na.rm = T)/sum(!is.na(baseline$cb_I))*100, 2)`\% and `r round(sum(endline$cb_I>0, na.rm = T)/sum(!is.na(endline$cb_I))*100, 2)`\% of apprentices are estimated to generate a positive net benefit during training at baseline and endline, respectively; using Model II, `r round(sum(baseline$cb_II>0, na.rm = T)/sum(!is.na(baseline$cb_II))*100, 2)`\% and `r round(sum(endline$cb_II>0, na.rm = T)/sum(!is.na(endline$cb_II))*100, 2)`\% of apprentices are estimated to do so; using Model III, `r round(sum(baseline$cb_III>0, na.rm = T)/sum(!is.na(baseline$cb_III))*100, 2)`\% and `r round(sum(endline$cb_III>0, na.rm = T)/sum(!is.na(endline$cb_III))*100, 2)`\% of apprentices are estimated to do so; using Model IV, `r round(sum(baseline$cb_IV>0, na.rm = T)/sum(!is.na(baseline$cb_IV))*100, 2)`\% and `r round(sum(endline$cb_IV>0, na.rm = T)/sum(!is.na(endline$cb_IV))*100, 2)`\% of apprentices are estimated to do so; using Model V, `r round(sum(baseline$cb_V>0, na.rm = T)/sum(!is.na(baseline$cb_V))*100, 2)`\% and `r round(sum(endline$cb_V>0, na.rm = T)/sum(!is.na(endline$cb_V))*100, 2)`\% of apprentices are estimated to do so.

Further tabulations can be found in the Appendix: Table \@ref(tab:tbl-appnetbenefitsbywave) shows baseline and endline results, while Table \@ref(tab:tbl-appnetbenefitsbytrade) reports cost and benefit estimates by trade. 

<!--
Mean benefits per apprentice total `r mean(df$cb_I/605, na.rm = T)` \$US when using Model I, `r mean(df$cb_II/605, na.rm = T)` \$US using Model II, `r mean(df$cb_III/605, na.rm = T)` \$US using Model III, `r mean(df$cb_IV/605, na.rm = T)` \$US using Model IV, and `r mean(df$cb_V/605, na.rm = T)` \$US using Model V.

To account for outliers, we truncate estimated net benefits per apprentice at the 1st, 5th and 10th percentiles and present the truncated distributions in a histogram and kernel density plot, shown in Figure \@ref(fig-apphist) in the Appendix. Using Model I, mean net benefits amount to `r mean(df[between(df$cb_I, quantile(df$cb_I, .01, na.rm = T), quantile(df$cb_I, .99, na.rm = T)), ]$cb_I/605, na.rm = T)` \$US per apprentice when dropping the top and bottom percentile, `r mean(df[between(df$cb_I, quantile(df$cb_I, .05, na.rm = T), quantile(df$cb_I, .95, na.rm = T)), ]$cb_I/605, na.rm = T)` \$US per apprentice when the top and bottom 5% are dropped, and `r mean(df[between(df$cb_I, quantile(df$cb_I, .1, na.rm = T), quantile(df$cb_I, .9, na.rm = T)), ]$cb_I/605, na.rm = T)` \$US per apprentice when the top and bottom deciles are dropped. -->

<!-- Using Model II, mean net benefits per apprentice are `r mean(df[between(df$cb_II, quantile(df$cb_II, .01, na.rm = T), quantile(df$cb_II, .99, na.rm = T)), ]$cb_II/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df[between(df$cb_II, quantile(df$cb_II, .05, na.rm = T), quantile(df$cb_II, .95, na.rm = T)), ]$cb_II/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df[between(df$cb_II, quantile(df$cb_II, .1, na.rm = T), quantile(df$cb_II, .9, na.rm = T)), ]$cb_II/605, na.rm = T)` \$US when the top and bottom deciles are dropped. -->

<!-- Using Model III, mean net benefits per apprentice are `r mean(df[between(df$cb_III, quantile(df$cb_III, .01, na.rm = T), quantile(df$cb_III, .99, na.rm = T)), ]$cb_III/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df[between(df$cb_III, quantile(df$cb_III, .05, na.rm = T), quantile(df$cb_III, .95, na.rm = T)), ]$cb_III/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df[between(df$cb_III, quantile(df$cb_III, .1, na.rm = T), quantile(df$cb_III, .9, na.rm = T)), ]$cb_III/605, na.rm = T)` \$US when the top and bottom deciles are dropped. -->

<!-- Using Model IV, mean net benefits per apprentice are `r mean(df[between(df$cb_IV, quantile(df$cb_IV, .01, na.rm = T), quantile(df$cb_IV, .99, na.rm = T)), ]$cb_IV/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df[between(df$cb_IV, quantile(df$cb_IV, .05, na.rm = T), quantile(df$cb_IV, .95, na.rm = T)), ]$cb_IV/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df[between(df$cb_IV, quantile(df$cb_IV, .1, na.rm = T), quantile(df$cb_IV, .9, na.rm = T)), ]$cb_IV/605, na.rm = T)` \$US when the top and bottom deciles are dropped. -->

<!-- Using Model V, mean net benefits per apprentice are `r mean(df[between(df$cb_V, quantile(df$cb_V, .01, na.rm = T), quantile(df$cb_V, .99, na.rm = T)), ]$cb_V/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df[between(df$cb_V, quantile(df$cb_V, .05, na.rm = T), quantile(df$cb_V, .95, na.rm = T)), ]$cb_V/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df[between(df$cb_V, quantile(df$cb_V, .1, na.rm = T), quantile(df$cb_V, .9, na.rm = T)), ]$cb_V/605, na.rm = T)` \$US when the top and bottom deciles are dropped.

-->



<!---
The ITT results found in Table 10 estimate the effect of the training on hours worked in
paid labor (which includes any paid employment, including paid labor in agriculture); in selfemployment
(which includes both work on family-owned land and in self-owned or familyowned
business); and in skill development (such as school, job or trade training) during and
after training. Treatment assignment and training participation leads to very large increases
in time spent on human capital development (i.e. training) during the training period. Being
assigned to the treatment group leads to 262-351 extra hours of training. Since training in
most professions lasted over three months (the average training duration was 13-14 weeks),
this is a reasonable estimate, and suggests that the training kept all trainees quite busy over the
entire training period (between 14 and 27 hours per week, on average). This is particularly
true for males, partly due to more regular attendance (see Table 3).
Investing all this time in training displaced many hours of work in both paid labor and in
self-employment, but not in a one-to-one fashion. Using the results without baseline controls,
the hours spent in paid labor and self-employment declined by 26 and 40 hours for females and
32 and 96 hours for males, respectively. This suggests that a sizable number of the hours in
training come from displaced paid labor and self-employment hours: 21% for females and 37%
for males. Notably, this effect is statistically different for men and women. Thus, the
opportunity cost of attending the training in terms of both time and forgone earnings is
substantial and affect men and women in slightly different ways.
Turning our attention to the effects of treatment assignment on time use after the training
is completed, we find that the most salient consequence of the training program is continued
investment in human capital for men. Males increase total hours spent on skill development
(through school or other job training) by 22 hours (29 hours including baseline controls), or 11-
14.5 hours per month after the training is over. While we do not have the data to test the longrun
labor market effects of continued investment in human capital, it is possible that increased
investment in skill development may have significant and lasting implications for labor market
opportunities in the long run (Attanasio et al. 2015). On the other hand, we observe no such
effect for women suggesting that men and women react differently to training programs in
terms of continued human capital investment, even when controlling for occupation. In Table
3, we shows that trainees in male dominated occupations generally have a better experience
with the MC (they are more likely to receive food and stipend support, and to find paid work
from MC), which makes it possible that men have added opportunities to continue skill
development with their MCs after the actual training period concludes.
We show in Table 11 that this extra time spent on training comes at a financial cost to the
trainees, with different consequences by gender. There is no discernible impact of training on
an economic index we construct (a composite of last week’s total earnings and total monthly
expenditure) for either gender. However, participation in the training forces women to draw
--->

<!---
Averaging over both survey waves, the largest fees reported by the firms are training fees, totalling `r round(mean(df$fee_formation, na.rm = T), 0)` FCFA (`r round(mean(df$fee_formation/605, na.rm = T), 0)` \$US). The remaining fees total `r round(mean(df$fee_entry, na.rm = T) + mean(df$fee_liberation, na.rm = T) + mean(df$fee_materials, na.rm = T) + mean(df$fee_contract, na.rm = T) + mean(df$fee_application, na.rm = T), 0)` FCFA (`r round((mean(df$fee_entry, na.rm = T) + mean(df$fee_liberation, na.rm = T) + mean(df$fee_materials, na.rm = T) + mean(df$fee_contract, na.rm = T) + mean(df$fee_application, na.rm = T))/605, 0)` \$US).
--->

<!---
As argued by @muhlemann2016: "_Dual apprenticeship training requires a firm to make a training investment, mainly in the form of personnel costs for apprentice instructors and paying apprentices a salary. However, a training firm also receives a benefit because apprentices work productively during training._"

_Dual apprenticeship training is a special type of training, as apprentices are hired specifically for the period of the apprenticeship programme, and the apprenticeship contract expires automatically after the end of training. Lindley (1975) discusses the two main motives as to why a firm would be willing to take on apprentices: (i) production motive, and (ii) investment motive (see also Merrilees, 1983). For the first motive, the relevance of the costs and benefits of apprenticeship training are immediately clear: if a firm’s main motive is to use apprentices for production during the apprenticeship period and does not wish to subsequently retain the graduated apprentices, then a firm would never hire an apprentice if the expected training costs exceed the expected training benefits. A similar logic applies if the firm expects that most apprentices will leave the firm voluntarily after training, for example, to pursue further studies or to go on and work for a different firm. Consequently, when a firm trains according to the production motive, apprentice pay must adjust so that the productive contribution of the apprentice covers a firm’s training expenditures._


_@hauschildt2018: Cost-benefit calculations of in-company training are regularly conducted in Germany by the BIBB following the work of the Sachverständigenkommission Kosten und Finanzierung der beruflichen Bildung, better known as the Edding Commission, in 1974. Since 1980, studies have regularly been carried out (1991, 2007/2008, 2012/2013, and 2017/18) (see Wenzelmann, et al., 2015, and bibb.de12)...a total of 3,000 companies form the basis of calculations. The first results of a representative large-scale Swiss study (Schweri et al., 2003) were based on a survey of 4,778 companies. Switzerland is currently running its fourth cost-benefit study, based on 3,000 firms._

_The methodology behind the German and Swiss large-scale assessments is the same. Research in both projects is based on company interviews, conducted face-to-face or on the phone, whereas in the first rounds of data collection, paper and pencil methods were used – nowadays, data entry is done via computer. Both studies aim primarily at providing representative results that help to understand developments from a national perspective. Comparison projects between the two country’s initiatives are possible and also provided (see Pfeifer, Wenzelmann, and Wolter, 2016)._

_Another approach is offered by QEK/CBQ, a method established by the University of Bremen in Germany, that has been high-income as an online measurement instrument based on self-evaluation and designed to deliver individual, company-specific results. The initiative to set up this instrument was taken within a project called IBB 2010 (Innovative Berufliche Bildung 2010) that was launched in 2006 as part of a larger EU-project consisting of a total of 14 sub-projects. Aiming at proving training firms with a tool-kit to investigate an individual cost-benefit situation, but simultaneously opening the opportunity to consulting activities in this domain, the instrument was not meant to be applied on a very large-scale as it was /is the case with the original BIBB approach. Since 2006, the QEK instrument has been applied in a number of local and regional projects, for example in Bremen/Bremerhaven, or in Saxony, as well as in a national sector study in elderly care. To date, about 1,000 companies have applied QEK in Germany so far. In addition, it has been modified and adapted for South Africa, where approximately 150 companies have used it thus far._
--->


<!---
Raw materials take a particular toll on small firms (no more than one employee and one apprentice) and carpentry workshops, accounting for xx% of total costs for the former and xx% of total costs for the latter. Since questions addressed apprenticeship in the firm in general, not a specific program, we are not able to differentiate between CQP and traditional apprentices in estimating cost of training, and will assume equal costs in the net benefit analysis in the following section.
--->

<!-- \noindent An even more comprehensive accounting of firm costs and benefits from training may account for factors which are not directly observable, such as technological upgrading and savings on hiring costs, as well as benefits that materialize in the future, such as apprentice labor after graduation [@walther2007]. However, savings on hiring costs are bound to be minimal, as firms are composed primarily of apprentices, and future productive value can also be ignored at first pass, since most apprentices strike out on their own upon graduation. The costs of recruiting and hiring apprentices may be an important omitted cost which was not covered by the survey, but even in this regard, many firms reported that parents and apprentices approached them rather than _vice-versa_. -->

``` {r df1, include = FALSE}
df1 <- df %>% group_by(FS1.2, wave) %>% summarise_at(c("cb_I", "cb_II", "cb_III", "cb_IV", "cb_V", "FS6.1", "FS5.1", "FS4.7", "FS4.1", "profits"), mean) %>% ungroup() %>% mutate(cb_I = cb_I * FS6.1, cb_II = cb_II * FS6.1, cb_III = cb_III * FS6.1, cb_IV = cb_IV * FS6.1, cb_V = cb_V * FS6.1)
baseline1 <- df1 %>% filter(wave == 0) %>% select(contains("cb"), FS1.2, FS4.1, FS4.7, FS5.1, profits)
endline1 <- df1 %>% filter(wave == 1) %>% select(contains("cb"), FS1.2, FS4.1, FS4.7, FS5.1, profits)
```

``` {r tbl-cblong, echo = FALSE}
```

\noindent In Table \@ref(tbl-cblong), we also investigate the total net benefits from training accruing to the firm. We do this by averaging apprentice-level fee and allowance data for each firm and multiplying it by the total number of apprentices training in that firm (training costs were already reported at the firm level, while apprentice and trainer productivity estimates were based on firm-level wages and training schedules). We arrive at estimated firm-level net  benefits of training by combining the various categories of costs and benefits using Models I-V from before. 

Mean estimated benefits per firm average range from `r mean(df1$cb_II/605, na.rm = T)` \$US per year to `r mean(df1$cb_III/605, na.rm = T)` \$US per year, with the large variance in estimates driven once more by apprentice productivity estimates. Table \@ref(tbl-cblong) reports estimates by firm size, showing that the largest firms in the sample, through significantly higher reported wages, have significantly higher estimated apprentice productivity.

Allowances are, as in the individual-level estimations, by far the largest cost related to training. Firm-level aggregation suggests that our methodology may in fact overestimate apprentice allowances: for all but the smallest firms, total estimated allowances are on average higher than total firm expenditures reported by the firm. 

As with the individual-level estimates, the majority of firms are clustered around zero net benefits for all cost-benefit models, with long left and right tails depending on the model used. Models I and II exhibit long left tails, while models III-V have long right tails, albeit for fewer observations (plotted with their means in Figure \@ref(fig:fig-firmhist) in the Appendix). 

To generate a firm-level distribution of net benefits, apprentice benefits are averaged at the firm level. This firm-specific mean benefit is then multiplied by the number of apprentices training in the firm, thus assuming that the mean net benefits associated with apprentices for whom data was collected are, on average, equal to the net benefits for all unobserved apprentices in the same firm. Net firm benefits total, on average, `r mean(df1$cb_I/605, na.rm = T)` \$US using Model I, `r mean(df1$cb_II/605, na.rm = T)` \$US using Model II, `r mean(df1$cb_III/605, na.rm = T)` \$US using Model III, `r mean(df1$cb_IV/605, na.rm = T)` \$US using Model IV, and `r mean(df1$cb_V/605, na.rm = T)` \$US using Model V.

<!-- Truncating again at the 1st, 5th and 10th percentiles and present the truncated distributions in a histogram and kernel density plot, shown in Figure \@ref(fig:hist2) in the Appendix. Using Model I, mean net benefits amount to `r mean(df1[between(df1$cb_I, quantile(df1$cb_I, .01, na.rm = T), quantile(df1$cb_I, .99, na.rm = T)), ]$cb_I/605, na.rm = T)` \$US per apprentice when dropping the top and bottom percentile, `r mean(df1[between(df1$cb_I, quantile(df1$cb_I, .05, na.rm = T), quantile(df1$cb_I, .95, na.rm = T)), ]$cb_I/605, na.rm = T)` \$US per apprentice when the top and bottom 5% are dropped, and `r mean(df1[between(df1$cb_I, quantile(df1$cb_I, .1, na.rm = T), quantile(df1$cb_I, .9, na.rm = T)), ]$cb_I/605, na.rm = T)` \$US per apprentice when the top and bottom deciles are dropped. -->

<!-- Using Model II, mean net benefits per apprentice are `r mean(df1[between(df1$cb_II, quantile(df1$cb_II, .01, na.rm = T), quantile(df1$cb_II, .99, na.rm = T)), ]$cb_II/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df1[between(df1$cb_II, quantile(df1$cb_II, .05, na.rm = T), quantile(df1$cb_II, .95, na.rm = T)), ]$cb_II/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df1[between(df1$cb_II, quantile(df1$cb_II, .1, na.rm = T), quantile(df1$cb_II, .9, na.rm = T)), ]$cb_II/605, na.rm = T)` \$US when the top and bottom deciles are dropped. -->

<!-- Using Model III, mean net benefits per apprentice are `r mean(df1[between(df1$cb_III, quantile(df1$cb_III, .01, na.rm = T), quantile(df1$cb_III, .99, na.rm = T)), ]$cb_III/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df1[between(df1$cb_III, quantile(df1$cb_III, .05, na.rm = T), quantile(df1$cb_III, .95, na.rm = T)), ]$cb_III/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df1[between(df1$cb_III, quantile(df1$cb_III, .1, na.rm = T), quantile(df1$cb_III, .9, na.rm = T)), ]$cb_III/605, na.rm = T)` \$US when the top and bottom deciles are dropped. -->

<!-- Using Model IV, mean net benefits per apprentice are `r mean(df1[between(df1$cb_IV, quantile(df1$cb_IV, .01, na.rm = T), quantile(df1$cb_IV, .99, na.rm = T)), ]$cb_IV/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df1[between(df1$cb_IV, quantile(df1$cb_IV, .05, na.rm = T), quantile(df1$cb_IV, .95, na.rm = T)), ]$cb_IV/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df1[between(df1$cb_IV, quantile(df1$cb_IV, .1, na.rm = T), quantile(df1$cb_IV, .9, na.rm = T)), ]$cb_IV/605, na.rm = T)` \$US when the top and bottom deciles are dropped. -->

<!-- Using Model V, mean net benefits per apprentice are `r mean(df1[between(df1$cb_V, quantile(df1$cb_V, .01, na.rm = T), quantile(df1$cb_V, .99, na.rm = T)), ]$cb_V/605, na.rm = T)` \$US when dropping the top and bottom percentile, `r mean(df1[between(df1$cb_V, quantile(df1$cb_V, .05, na.rm = T), quantile(df1$cb_V, .95, na.rm = T)), ]$cb_V/605, na.rm = T)` \$US when the top and bottom 5% are dropped, and `r mean(df1[between(df1$cb_V, quantile(df1$cb_V, .1, na.rm = T), quantile(df1$cb_V, .9, na.rm = T)), ]$cb_V/605, na.rm = T)` \$US when the top and bottom deciles are dropped. -->

According to Model I, `r round(sum(baseline1$cb_I>0, na.rm = T, na.rm = T)/count(baseline1)*100, 2)`\% and `r round(sum(endline1$cb_I>0, na.rm = T)/count(endline1)*100, 2)`\% of firms are estimated to earn a positive net benefit from training (positive net benefits) at baseline and endline, respectively; using Model II, `r round(sum(baseline1$cb_II>0, na.rm = T)/count(baseline1)*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_II>0, na.rm = T)/count(endline1)*100, 2)`\% at endline, using Model III, `r round(sum(baseline1$cb_III>0, na.rm = T)/count(baseline1)*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_III>0, na.rm = T)/count(endline1)*100, 2)`\% at endline; using Model IV, `r round(sum(baseline1$cb_IV>0, na.rm = T)/count(baseline1)*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_IV>0, na.rm = T)/count(endline1)*100, 2)`\% at endline; using Model V, `r round(sum(baseline1$cb_V>0, na.rm = T)/count(baseline1)*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_V>0, na.rm = T)/count(endline1)*100, 2)`\% at endline.
 
Apprentice training represents a relatively large fraction of firm costs, or revenues, depending on the preferred model. At the upper extreme (Model II, untruncated), the average ratio of total net training benefits to total firm revenues is `r round(baseline1 %>% filter(cb_V >= 0, FS4.7 > 0, FS4.1 > 0) %>% mutate(ratio = cb_V/FS4.7/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)` for firms that benefit from or break even on training, and the ratio of total (negative) net benefits from training to total firm expenses is `r round(baseline1 %>% filter(cb_V < 0, FS5.1 > 0, FS4.1 > 0) %>% mutate(ratio = -cb_V/FS5.1/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)`. At the low-benefit extreme (Model II), the average ratio of total net training benefits to total firm revenues is `r round(baseline1 %>% filter(cb_II >= 0, FS4.7 > 0, FS4.1 > 0) %>% mutate(ratio = cb_II/FS4.7/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)` for firms that benefit from or break even on training, and the ratio of total (negative) net benefits from training to total firm expenses is `r round(baseline1 %>% filter(cb_II < 0, FS5.1 > 0, FS4.1 > 0) %>% mutate(ratio = -cb_II/FS5.1/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)`.

Using calculated instead of reported profits, the average ratio of total net training benefits (Model II) to total firm revenues is `r round(baseline1 %>% filter(cb_V >= 0, profits > 0, FS4.1 > 0) %>% mutate(ratio = cb_V/profits/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)` for firms that benefit from or break even on training, and the ratio of total (negative) net benefits from training to total firm expenses is `r round(baseline1 %>% filter(cb_V < 0, FS5.1 > 0, FS4.1 > 0) %>% mutate(ratio = -cb_V/FS5.1/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)`. At the low-benefit extreme (Model II), the average ratio of total net training benefits to total firm revenues is `r round(baseline1 %>% filter(cb_II >= 0, profits > 0, FS4.1 > 0) %>% mutate(ratio = cb_II/profits/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)` for firms that benefit from or break even on training, and the ratio of total (negative) net benefits from training to total firm expenses is `r round(baseline1 %>% filter(cb_II < 0, FS5.1 > 0, FS4.1 > 0) %>% mutate(ratio = -cb_II/FS5.1/FS4.1) %>% summarise_at("ratio", mean, na.rm = T), 2)`.



<!---
In contrast to dual apprenticeship in the European context, in which apprentice wages represent the bulk of training costs, most apprentices in Bénin receive no regular wages whatsoever (see Table \@ref(tab:wages)). Allowances or "pocket money" for minor expenditures such as food and transportation disbursed by training firms amount to a small percentage of total training costs. On the other hand, apprenticeship fees, paid by apprentices directly to the master trainer, contribute significantly more to total firm revenues than in Swiss or German firms.

This implies that for firms that benefits from apprenticeship, apprenticeship fees and apprentice productivity are estimated to only contribute slightly to their total revenues. For firms that do not turn a profit on their apprentices, on the other hand, estimated training costs are actually higher on average than the total costs reported by firm owners.
--->

``` {r tbl-firmregs, results = 'asis'}
```


\noindent In addition to direct financial benefits associated with training, which are reflected by a positive balance in the net benefit calculations presented above, apprenticeship training may affect firm productivity through a variety of additional channels. For the CQP program in particular, participating apprentices may acquire skills at a faster pace than their traditional counterparts as a direct result of their theoretical training. Moreover, theoretical training may "spill over" to the master trainer and other employees in the workshop, for instance by introducing new technologies or improving knowledge about the operation of existing workshop machinery. Additional apprentices in general may improve firm productivity by encouraging the owner to hire more employees (e.g. as trainers) or through investments in additional machinery. Indeed, evidence from previous studies indicates that small firms in Uganda and Ghana, when randomly assigned apprentices to train, increased profits by up to 15\% per apprentice [@hardy2022; @alfonsi2020]. 

Pooled OLS estimation results are shown in Columns 1, 3 and 5 of Table \@ref(tab:tbl-firmregs), while firm fixed effects are included in the specification shown in Columns 2, 4, and 6. Reported firm revenues increase by close to 50\% between the two waves, but are offset by rising costs and wages, to the point of eliminating any observed growth in profitability. According to the pooled specifications, additional apprentices are associated with smaller gains in revenue than the hiring of other types of workers, but higher profits (about 92 \$US annual profit per additional apprentice). Specifications including firm fixed effects to control for heterogeneity between firms, on the other hand, suggest that the hiring of non-apprentice employees is associated with both higher revenues and profits, while the number of apprentices is not a significant predictor of either. Regressions (5) and (6) reinforce the notion that firms hire more apprentices than non-apprentice employees: taking into account firm fixed effects, a single additional employee is associated with over three additional apprentice hires, on average. Finally, we note that the number apprentices from the 2019 cohort of the CQP program have no discernible association with firm revenues or profits.

# Conclusion {#conclusion}

This paper analyses the costs and benefits of apprenticeship training with informal firms in Bénin, with a particular focus on the 2019 cohort of apprentices participating in the CQP dual training program. Apprentice-firm data is used to (1) estimate the human capital gains accumulated by apprentices over the three-year observation period, (2) calculate the total benefits (primarily in the form of apprenticeship fees received) and (3) costs of training reported by firms, (4) calculate the net benefits accruing to firms using two models, one estimating apprentice contribution and foregone trainer productivity and one not, and (5) identify the association between apprenticeship training, and participation in the CQP program in particular, on firm profitability, apprentice human capital gains, and the net benefits accruing to firms per apprentice.

First, we find that there is visible selection into the CQP program. CQP applicants are better educated, older, and have spent more time as apprentices than other apprentices in the firm. Firms sending apprentices to the CQP have fewer than 10 total workers on average, of whom the majority are apprentices. 

Human capital gains as measured by trainers' subjective appraisal of apprentice experience and competence were significant, with improvements of .73 and .66 standard deviation between survey waves, respectively. Gains were concentrated in the masonry and carpentry sector, in larger firms, and among apprentices who did not apply for the CQP program.

Two cost benefit models help us understand how net benefits are distributed across firms and apprentice types. When only fees, allowances, and reported training costs are considered, firms suffer a net loss of `r mean(-df$cb_II/605, na.rm = T)` \$US per apprentice per year, and a total of `r mean(-df1$cb_II/605, na.rm = T)` \$US when accounting for the number of apprentices trained. Using reported wages for skilled and unskilled workers to estimate apprentice productive contributions and foregone trainer productivity reduces the net estimated losses, to `r mean(-df$cb_V/605, na.rm = T)` \$US per apprentice per year, and even a net benefit of `r mean(df1$cb_V/605, na.rm = T)` \$US per firm when projecting total benefits at the firm level. Apprentices who do _not_ apply to the CQP program are associated with somewhat lower net losses to their training firms when estimated using the first model.

Regression analysis reveals a stronger relationship between firm size (sans apprentices) and firm profits than exists for apprentice hires (CQP and otherwise). Similarly, apprentices hired at larger firms are associated with higher net benefits. There are no significant differences in effects observed for CQP applicants or CQP participants, though firms appear to select less experienced apprentices to apply for the program.

Interruptions due to the Covid-19 pandemic may have disrupted training in the period of observation: `r prop.table(table(df$COV15))[1]*100`% of apprentices reported reduced hours of training, while `r prop.table(table(df$COV15))[2]*100`% reported complete work stoppages at their training firm, due to the pandemic. Covid-19 is also likely to have had an outsized impact on participants in the CQP program, `r prop.table(table(df$COV15))[3]*100`% of whom reported that their training center had suspended classes. Among apprentices who participated in any external training, `r sum(df$YS4.44[!is.na(df$YS4.44)]>6)/length(df$YS4.44[!is.na(df$YS4.44)])*100`% reported working spending at least 10 days in this training in the preceding three months (approximately equal for the CQP subsample), while the average reported training duration was `r mean(df$YS4.46, na.rm = T)` hours. Thus, CQP apprentices reported training externally at only a marginally higher rate than apprentices who applied but were not accepted into the program. 

In sum, we find that training with informal firms fulfills its promise to apprentices, granting them the experience and competence to transition to self-employment upon graduation. Not all firms benefit from training on the balance, with about 60\% of firms suffering net losses from training and training costs generally outpacing the apprenticeship fees paid. Dual training in the form of the CQP program generates few observable benefits, either in terms of apprentice progression or benefits to the training firm. However, the timing of the current CQP cohort, being interrupted in the middle of the training schedule due to the Covid-19 pandemic, may have contributed to lower-than-usual outcomes at both the apprentice and firm level. 


<!---
Nevertheless, this paper finds little support for or against the continuation of the program. Given its difficulties with sustainable financing, and in light of criticisms centered on the administration of training centers and qualifications of classroom teachers, helping individual firms defray the high costs of traditional training, and focusing on material and equipment provision for traditional on-the-job apprenticeships, may be a better use of public financing in the short term. For the time being, in any case, dual training only represents a minuscule fraction of all apprenticeship training in the country of Bénin and across West Africa; with the slow growth of formal TVET programs, traditional apprenticeships are likely to remain a promising route for young Africans who lack the financial means or interest to pursue formal education.

_@mohrenweiser2009: The German dual apprenticeship system is often regarded as a potential model for other countries because it allows enterprises to provide and pay for training in specific and general human capital (Harhoff and Kane, 1997; Acemoglu and Pischke, 1999b; Steedman 2001)...We show that employing apprentices in trade, commercial, craft, and construction occupations instead of unskilled or semi-skilled employees has a positive impact on contemporary establishment performance. In contrast, apprentices in manufacturing occupations have higher relative training costs in comparison to their benefits — contemporary establishment performance declines when their share of employees increases while those of unskilled or semi-skilled employees decreases...The net costs assumption stems from a series of descriptive cost–benefit-studies conducted by the BIBB (Bardeleben et al., 1997; Beicht et al., 2004; Wenzelmann et al., 2009). These evaluations calculate relevant costs during apprenticeship training for about 50 occupations and compare them with the economic value of the productivity contribution of a typical apprentice. Evaluations are based on surveys of about 2500 (personnel) managers who assessed one occupation each. They consistently find that all occupations and almost all companies face sizeable net costs during the apprenticeship period...Fougère and Schwerdt (2002) find a positive effect of apprentices on value added only in medium-sized German establishments. Askilden and Nilsen (2005) analyse the recruitment of apprentices during the business cycle in Norway. They find that apprentices are substitutes for skilled workers and are primarily recruited in boom phases. Zwick (2007) finds on the basis of panel estimations that an increase in the share of apprentices does not decrease profits on average_
 

In the German and Swiss context, the primary benefit of training for firms is the productive value of apprentices during and after their apprenticeship. This value is generated when the apprentices is working in the firm, performing skilled or unskilled tasks that would otherwise be carried out by the firm owner, paid employees, or unpaid labor. These benefits accumulate after graduation if the apprentice is hired for less than the labor market wage for workers of equal productivity, as well as any savings on recruitment and hiring costs [@acemogluWhyFirmsTrain1998; @acemogluStructureWagesInvestment1999; @acemogluTrainingImperfectLabour1999].

In comparison to Swiss or German apprentices, youth in SSA are less likely to continue working for their training firm upon graduation. For example, in Ghana, three out of every four apprentices plan to start their own business within five years [@breyer2007; @ahadzie2009]. Many who don't set off on their own are unable to, citing "high start-up costs, the unavailability of land and hostile planning policies tend to obstruct the ability of newly qualified craftsmen to set up on their own" [@ahadzie2009]. Informal firms must thus recoup their costs during the training period (or within a few years) if hiring apprentices is to remain a profitable enterprise. Moreover, workshops rarely hire skilled wage employees, and thus any benefits from "screening" apprentices for their productive capabilities (i.e. wage compression) or savings in hiring costs play a relatively minor role in the training decision (Table \@ref(tab:tbl-desc)).

If this is the case, why do craftsmen in Bénin train at all? First, apprentices are the cheapest and most abundant source of labor, and can be hired to complete unskilled tasks at little overhead. Second, if trades can be mastered relatively quickly, apprentices will provide skilled labor in the later years of their graduation (and after graduation as well, for as long as they need to work to pay off outstanding training fees or save up for their own workshop). Finally, the apprentices (or, more often, their families) traditionally pay fees to the master craftsmen in return for the provision of training [@velenchik1995; @frazer2006; @bankole2020].


Moreover, recall that the total cost of a four-year CQP training program was estimated to be between \$165 and \$413 in 2012, of which the trainer is covers 10\%. Assuming the most conservative estimate of net benefits (Model II), the average firm owner accrues a net benefit of 38.64 \$US from training a CQP apprentice for four years, about covering the firms 10\% share of the most expensive cost estimate, $413. Assuming Model I and \$413 CQP cost, the training firm nearly covers CQP costs within a single year; when estimated apprentice productivity is included (Model II), firms recover CQP fees nine times over.

Finally, how do these numbers compare to evidence from high-income countries? Due to the disparities in firm size and productivity of informal firms in Bénin and training firms in Germany and Switzerland, where the majority of studies have been conducted, an informative comparison is difficult to make. Though cost-benefit studies from these countries do not report earnings data of training firms, they do suggest that whether firms recoup the costs of training depends on the firm, trade, and region. For instance, @hauschildt2018 reports that in Germany, productive contributions of apprentices covered about 70 percent of a company's training costs, while in Switzerland, net benefits per apprentice per year amounted to about \euro 2,500 [@strupler2012]. Without a clearer understanding of future benefits and better instruments to measure apprentice productivity, the true benefit of apprenticeship training in informal firms will remain ambiguous.

- _The number of trades in which the CQP and CQM are awarded must be increased. Indeed, while the Bénin craft trades register lists 310 trades, the CQP only covers 13 and the CQM fewer than 100. It is essential to develop the design tools needed to organize the CQP and CQM. This requires the training of a sufficient number of methodology experts across all departments nationwide. In addition to these methodology experts, trainers in all crafts and trades need to be trained in sufficient numbers._
- _There are no qualifications that recognize the learning done by employers. As a result, some employers still refuse to enter their apprentices for the CQP or the CQM, as they think that their apprentices will become better skilled than them. In order to maintain the system, it is essential to introduce a mechanism for validating and recognizing the skills of employers._
- _Both the apprentices and their employers show more interest in the CQM than the CQP, as the conditions for the CQP – namely required educational level (completed primary education), attendance at a (dual) training centre, preparation for and successful completion of an admission test, and the irregular financing of the FODEFCA – constitute obstacles that they cannot overcome. If these conditions are not reviewed or improved, there is a risk that the CQP will lose its candidates._
- _Internal and external financial resources must be mobilized to manage the increased number of trades and apprentices involved in the CQP and CQM._
- _The DEC needs to allocate or obtain timely budgetary resources to organize the CQM examinations._
- _UCIM-B demands responsibility for organizing the CQP and CQM examinations in accordance with the West African Economic and Monetary Union (WAEMU) community code on crafts. Dialogue between this body and the ministry responsible for training should facilitate resolving the matter of the transfer of powers_ [@ilo2021]
--->


\clearpage

# References {.unnumbered}
\noindent

<div id="refs"></div>

# (APPENDIX) Appendix {-} 

# Appendix A {-#appendix-a}

```{r child = "appendix-a.Rmd"}
```

# Appendix B {-#appendix-b}

\setcounter{figure}{0}
\renewcommand{\thefigure}{B\arabic{section}.\arabic{figure}}

<!-- 
```{r child = "appendix-b.Rmd"}
``` 
-->

