---
title: Costs and Benefits of the Dual System in the Context of Traditional Apprenticeship in Bénin
author: 
  - name: Bart Kudrzycki
    affiliation: Development Economics Group, ETH Zurich, Switzerland, <bartlomiej.kudrzycki@nadel.ethz.ch>
header-includes:
  - \usepackage{setspace}
date: \today
abstract: | 
  Traditional apprenticeships are an important source of skills for early school leavers in developing countries. Usually private arrangements between parents and informal firms, apprenticeship quality and costs are not subject to oversight or regulation, and the various transactions taking place between apprentice and firm have not been carefully documented to date. This paper sheds light on the institution of traditional apprenticeship through a study of a national apprenticeship reform called the CQP, which combines in-firm training with a weekly classroom component. Using two waves of surveys with firm owners and apprentices, we analyze the human capital gains and material costs and benefits associated with training. We find that the program did not generate improvements in apprentice skill or in firm profitability, possibly due to a strong selection of older, overqualified candidates into the program. We do find that traditional apprenticeship on the whole improves apprentices' sector-specific knowledge and experience. Fees paid by apprentices are found to be commensurate with the allowances they receive from the trainers, contradicting earlier work which suggest that firms use fees for financing. Estimates suggest that apprentice productivity must be substantial for firms to break even on training. The high proportion of apprentices as a fraction of the total workforce in firms, combined with the tendency for apprentices to leave their training workshop after graduation, suggests that apprentice productive contributions to the training firms are indeed substantial.

# This paper evaluates the impact of and provides a detailed account of the costs and benefits associated with (dual) apprenticeship training in the informal sector in Bénin using unique matched apprentice-firm survey data. To study the effectiveness of dual training in an informal economy, successful applicants to the _Certificat de Qualification Professionelle_, program, a national training program that combines in-firm training with classroom teaching, are compared to unsuccessful applicants and non-applicants three years after program start.  Human capital gains among all apprentices, as measured by trainers' subjective appraisal of apprentice experience and competence, are significant, with improvements of .73 and .66 standard deviation between survey waves, respectively. Gains were 3concentrated in the masonry and carpentry sector, in larger firms, and among apprentices who did not apply for the CQP program. To study net benefits of training for firms, two cost and benefit models are used. When only fees, allowances, and reported training costs are considered, firms suffer a net loss of 130 $US per apprentice per year: an average of 630 $US total per firm. When projected apprentice contributions and foregone trainer productivity are included, net estimated losses are reduced to 33 \$US per apprentice per year, and an overall estimated benefit of 218 $US per firm. Fees paid to the training firms account are not enough to defray the expenses associated with training, which are dominated by allowances paid to apprentices. Apprentices who do not apply to the CQP program are associated with somewhat lower net losses to their training firms when estimated using the first model. Apprentices hired at larger firms are associated with higher net benefits. In sum, we find that participation in the CQP program is not associated in any benefits for either trainees or firms relative to non-participating apprentices, though effectiveness of dual training may have been negatively impacted by the Covid-19 crisis.
  
thanks: |
  Thanks to Dario for the Markdown template.
keywords: [Informal labor markets, Dual training, Apprenticeship]
jel: [I26]
bibliography: cqp.bib
csl: 'apa-no-ampersand.csl'
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    template: template.tex
    extra_dependencies: ["flafter", "placeins"]
    fig_caption: yes
---

```{r cqp_setup, include=FALSE}
packages <- c("tinytex", "lfe", "lme4", "plm", "bookdown", "gtsummary", "labelled", "grid", "gridExtra", "rstatix", "kableExtra", "stargazer", "tidyverse")

options(repos = c(CRAN = "http://cran.rstudio.com"))

# install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages], silent = TRUE)
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(installed_packages)

# load functions
source("../functions/add_by_n.R")
source("../functions/detach_packages.R")

# load data
# source("../code/master.R") # to rerun all code periodically
load("../data/df.rda")
selected <- df %>% filter(SELECTED == 1)
baseline <- df %>% filter(wave == 0)
endline <- df %>% filter (wave == 1)
firms <- df %>% select(FS1.2, wave, FS3.4, firm_size, FS6.1, contains("cb")) %>% group_by(FS1.2, wave) %>% summarise_all(mean, na.rm = T) %>% ungroup()

# load code chunks
knitr::read_chunk('../code/tbls_body.R')
knitr::read_chunk('../code/tbls_appendix.R')
knitr::read_chunk('../code/figures.R')

# set R options
ggplot2::theme_set(ggplot2::theme_bw())
options(scipen=2, digits=2)

# set knitr options
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      echo = FALSE,
                      cache = TRUE,
                      fig.path='figures/',
                      out.extra = "",
                      floatfoot.envs = c("figure", "table", "longtabu"))
```


\doublespacing

# Introduction {#intro}


``` {r df1, include = FALSE}
df1 <- df %>% group_by(FS1.2, wave) %>% summarise_at(c("cb_I", "cb_II", "cb_III", "cb_IV", "cb_V", "FS6.1", "FS5.1", "FS4.7", "FS4.1", "profits"), mean) %>% ungroup() %>% mutate(cb_I = cb_I * FS6.1, cb_II = cb_II * FS6.1, cb_III = cb_III * FS6.1, cb_IV = cb_IV * FS6.1, cb_V = cb_V * FS6.1)
baseline1 <- df1 %>% filter(wave == 0) %>% select(contains("cb"), FS1.2, FS4.1, FS4.7, FS5.1, profits)
endline1 <- df1 %>% filter(wave == 1) %>% select(contains("cb"), FS1.2, FS4.1, FS4.7, FS5.1, profits)
```

In sub-Saharan Africa (SSA), interest in apprenticeships is on the rise. In countries with largely informal economies, traditional apprenticeships (also referred to as apprenticeships in the informal sector or informal apprenticeships) are one of the most important sources of skills for early school leavers, accounting for as much as 80 percent of technical and vocational training (TVET)\footnote{According to a survey of five countries, 20 percent of youth aged 25-34 had participated in an apprenticeship in the past, though participation varied by country and was as high as 35 percent in Ghana; in contrast, only 1 percent were enrolled in formal TVET, and about 9 percent in tertiary education (Filmer and Fox, 2014). A more recent estimate of enrollment in formal TVET is 6 percent across SSA (Hofmann et al., 2022)} and for as much as 90 percent of total employment in the crafts sector [@walther2007; @adams2013; @filmer2014; @worldbank2018]. As increasing numbers of youth in SSA suffer from a lack of labor attachment, underemployment, and poverty, informal sector training is seen by many policy experts as an important tool for tackling the youth employment challenge.

In contrast to formal TVET, which usually takes place exclusively in the classroom, traditional apprentices train on-the-job in informal sector microenterprises or workshops. Traditional apprenticeships involve a private contractual arrangement between an aspiring apprentice --- usually a school leaver between the age of 14 and 18 --- or his or her parents, and a master craftsman (MC) who agrees to train the apprentice for a duration of about three to four years for a fee [@bas1989]. Upon completion of the apprenticeship, the MC usually issues a certificate acknowledging the training; some apprentices continue to work for the same or for another workshop wage employees, though most seek to start their own business given access to sufficient capital [@frazer2006]. While unregulated at the national level, informal apprenticeships are nevertheless structured according to the dictates of tradition and the customs of local professional associations, and, in the context of highly informal economies, are generally considered to be more effective than formal TVET at delivering the skills demanded by the labor market [@ahadzie2009].

Despite their attractiveness as a source of skills for young school leavers, the unregulated nature of traditional apprenticeships also gives rise to a number of potential market failures that may negatively affect their provision and have led to calls for their reform [@walther2011]. For instance, in the absence of complete, enforceable contracts, firms may be unable to commit to providing general skills training [@acemoglu1998; @acemoglu1999; @dustmann2012]. Apprentice productivity may also be so low that subsistence levels (paid in the form of "chop money" by the firm owner) are greater than the returns to training, resulting in its under-provision. Firms may also fear "poaching" of newly trained apprentices by competitors --- a particularly salient problem for small enterprises, which experience higher employee turnover and offer fewer opportunities for career advancement [@mcintosh2011] --- though evidence of this is limited for the African context.

Quality is also affected by the unregulated nature of informal apprenticeships, and may have adverse consequences for participating youth: apprentices may be exposed to inexperienced trainers who keep them in their apprenticeship for too long [@bas1989], or experience limited labor market mobility into formal sector wage jobs due to the lack of formal accreditation systems [@acemoglu2000; @worldbank2018; @alfonsi2020]. Policy makers have been increasing interested in reforms to address such market failures. such as the competence-based, nationally-accredited certification of informal apprenticeship which has recently been introduced in countries including Niger, Togo, Malawi and Tanzania.

An alternative reform is the introduction of a classroom component to an otherwise traditional apprenticeship, producing a hybrid "dual system" comparable to that of the Swiss and German variety [@walther2011]. Dual systems promise to increase training quality by introducing a state-regulated classroom component, while also improving the signalling ability of apprentices by offering official, nationally-recognized certification. In SSA, dual apprenticeship schemes were first introduced in Bénin and Togo in the 1980s by the Hans Seidel Foundation, a German NGO, and apprenticeship reforms based on the dual system have since been introduced in Mali, Côte d'Ivoire, Senegal, Tanzania, Togo, and Niger [@walther2011; @ilo2020a]. Many of these schemes have struggled with financing, MC engagement, and integration into the existing national TVET and regulatory frameworks; nevertheless, with its potential to simultaneously harness the abundance of training firms in the informal sector and the growing demand among parents and youth for formal education, dual system formalization remains a promising approach to TVET reform.

In this paper, we analyze the impact of a national dual system program on participating apprentices and firms in Bénin. The program, called _Certificat de Qualification Professionnelle_ (CQP), appends weekly classroom training at a local training center to otherwise traditional in-firm apprenticeship training in an informal firm. Using two waves of apprentice-firm survey data collected for 427 apprentices training in 197 firms, we assess the impact of this weekly classroom training component on learning outcomes, and estimate the marginal effect of apprentice participation in dual system training on firm size and profits.

Studies of vocational training interventions combining on-the-job and classroom teaching in middle-income countries have reported modest yet persistent increases in earnings together with mixed impacts on employment [@card2011; @attanasio2011; @ibarraran2014; @alzua2016;  @attanasio2017; @ibarraran2019]. Similar interventions in LICs have been characterized by low take-up, high dropout, and low efficacy [see @blattman2015; @ghisletta2021; @tripney2013 for an overview]. These programs tend to be shorter than the one studied in this paper (up to several months, rather than years), and focus on employment in the formal rather than the informal sector. To our knowledge, only one paper has attempted to quantify the impact of dual-system training in Sub-Saharan Africa: in a randomized experiment in Côte d'Ivoire, @crepon2019 found that youth offered a stipend for an apprenticeship that combined 12 to 24 months of on-the-job training with theoretical classes at local training institutions earned 15 percent more after three years, were involved in more complex and non-routine tasks\footnote{Enrollment in subsidized dual apprenticeship training increased the likelihood of undertaking non-routine analytical tasks by .24 standard deviations (SDs) and non-routine interpersonal tasks by 0.08 SDs relative to non-treated traditional apprentices. A task intesity index was found to be .21 SDs lower for dual apprentices, suggesting that dual apprentices were involved in a wider range of tasks.}, and received training certification at a higher rate than non-treated youth. We study a similarly structured dual training program, but one that is about twice as long and does not involve any direct subsidies or eliminate fees.

A related literature studies the incentive structure behind the dual system by analyzing the costs and benefits of apprenticeship for the training firm. A number of such studies have been conducted in the European context with the aid of surveys and simulations [see, e.g., @muhlemann2016; @muhlemann2019; @muhlemann2014], but have only recently begun to generate interest in lower-middle and low-income countries. Examples include @bolli2021, who find that training costs outweigh benefits in Serbia (with larger firms suffering smaller losses), and @bolli2020, who show that training firms in Nepal generally profit from training, with little variation in net benefits across firm size. To our knowledge, ours is the first rigorous cost-benefit study of dual training conducted in SSA.

Though traditional apprenticeships are very common in West Africa and across SSA, there is limited direct empirical evidence of their impact on the skills or labor market outcomes of apprentices. Long-term returns to informal training have been shown to be quite heterogeneous in Ghana, benefiting youth with lower levels of education the most [@monk2008]. An experimental study in Uganda found that six months of in-firm training measurably improved apprentices' skills, and that these skills persisted two to three years after the end of training [@alfonsi2020]. However, skills acquired in informal training tend to be firm-specific, and thus more likely to lead to self-employment than to quick career progression in the formal sector than formal schooling [@frazer2006; @hardy2019; @alfonsi2020].

Studies from SSA also suggest that informal apprenticeship training tends to have positive effects on microenterprise growth and profitability. Using data on formal manufacturing firms from Kenya, Zimbabwe and Ghana, @rosholm2007 observed a significant wage increase (of about 20%) in firms that trained in the previous 12 months, with large firms benefiting more than small ones. @hardy2022 found that randomly assigning an apprentice to informal firms in Ghana increased firm size by about half a worker, and firm revenues by 5-15 percent per apprentice. While @crepon2019 look at the impact of fee subsidies on firms' apprentice and employee stocks, they did not estimate the change in size or revenues that a firm can expect from hiring additional apprentices. Our study is thus the first, to our knowledge, to report the impact of dual system training in SSA on firm size and profits.

Finally, a number of studies have examined the financial arrangements between traditional apprentices and informal firms. @velenchik1995 studies the structure of apprenticeship contracts in small informal firms in Ghana, identifies three main transactions between apprentice and firm --- apprentice wages, fees and allowances --- and distinguishes between two broad types of contracts, namely those with and those without training fees. She finds that firms that do not charge fees are smaller and tend to offer more specific training. @velenchik1995 and @frazer2006 also suggest that training fees may be a substantial source of financing for some firms, but do not provide estimates of the allowances, wages and other training costs that these fees are meant to offset. This study attempts to fill this gap.

We find that, in general, all apprentices gain trade-specific human capital over the three observed years of training. However, we are unable to show that participation in dual system training contributes to additional learning. Youth with low learning scores at baseline make the largest gains in human capital over the course of the study; since we observe that MCs send older, more educated and more experienced apprentices to apply for the CQP program, this could explain the low level of efficacy observed for dual training.

In terms of financial arrangements, we find that fees paid by apprentices are not enough to cover the costs of their training in general, and the regular allowances they receive from their MC in lieu of wages in particular. Though estimating apprentice productivity presents a challenge with the available data, we find that between `r round(sum(baseline1$cb_V>0, na.rm = T)/sum(!is.na(baseline1$cb_V))*100, 0)` and `r round(sum(baseline1$cb_IV>0, na.rm = T)/sum(!is.na(baseline1$cb_IV))*100, 0)` percent of firms benefit from their training activities when apprentice productivity is accounted for (the vast majoritiy face net losses when it is not). While larger firms generate positive net benefits from training on average, smaller firms consistently report losses.

<!-- Only 33%-45% of apprentices are estimated to generate a profit for their MC; a similar range captures the proportion of firms that are estimated to profit (generate a positive net benefit) from their training activities. Finally, we find that larger firms benefit the most, while smaller firms suffer the largest losses, suggesting that economies of scale contribute to the incentive structures surrounding firm training. -->

The paper proceeds as follows. Informal apprenticeship in Bénin, the CQP program, and the survey data used for the analysis are presented in Section \@ref(data). Results are presented in Section \@ref(cqp_results). Section \@ref(conclusion) concludes.


<!-- @bolli2021  _participating companies are relatively large compared to the large share of micro and small companies in the Nepali economy, employ a relatively high share of high-skilled workers and face high labor market tightness....companies make average net benefits of about 17’000 NPR [137 USD - over 20 months]. These net benefits are substantially higher for the electrical engineering occupation (62’000 NPR) than for the mechanical engineering occupation (6’000 NPR), mainly due to the higher productivity of apprentices from the beginning onwards. The results suggest that apprentices create 270’000 NPR in productive value in the 20 months of workplace training. These benefits are reduced substantially by the costs that amount to 252’000 NPR. The wages of apprentices represent the most important cost component (56%), 9 Note that total benefits and instructor costs are calculated as the surfaces shown in Figure 6 and Figure 7. To derive total benefits and instructor costs based on values shown in Figure 8, it is important to remember that Figure 8 shows values after some time has elapsed rather than averages of a month. Hence, calculating total benefits and instructor costs based on these values consists of adding up the values but give a weight of 50% to the observations after three and 23 months. 25 followed by instructor costs (35%). Non-wage compensation is the least important cost component (9%). However, total benefits are larger than total costs, suggesting that the difference yields a net benefit for companies of 17’000 NPR._ -->

<!-- @deamesti2021 _by participating in dual training, the surveyed companies incur gross costs of $3,298,440 Chilean pesos (CLP) per year per student on average. This corresponds to US$4,650 or about 4.3 monthly wages of a skilled employee in the same area as the dual student. These gross costs consider four different types of costs: (1) training costs (57%), (2) student labour costs (28%), (3) administrative costs (13%), and (4) educational supplies (2%). Regarding the benefits, companies capture around $1,180,207 CLP per year per student. The benefits come from three sources: (1) the productive value that students contribute through their skilled and semi-skilled activities (79%), (2) savings in hiring (2%), and (3) savings in adjustment or induction (19%)._ -->

<!-- @chen2013 _have analysed the costs and benefits of apprenticeship training in a cooperation project between the Guangzhou Communications Technician Institute and Guangzhou Machine Tool Works Co, Ltd. using a calculation methodology also based on CBQ/QEK, but -- with regard to the cost benefit section -- adapted to the Chinese context. Data were collected within the framework of a three-year apprenticeship programme with 28 apprentices taking part. According to this field study, a company was able to obtain net benefits in each year of training, the first year being the most beneficial (for the exact figures, see Appendix, table A.5). This was due to the fact that leaners only received very little or sometimes even no training allowances in the first year of training. Moreover, it was discovered that the share of training allowances was about 90 per cent of the total cost of training, which easily explained the rather unusual result of high benefits in a first year of training in China._ -->


<!-- Dual "apprenticeship is a form of skill acquisition in the context of production, or on-the-job training. The apprentice, in addition to learning a trade, provides labor services to the firm. The relationship between the apprentice and the firm, therefore, includes two types of transactions, one being the provision of labor services by the apprentice to the firm, the other being the provision of training services by the firm to the apprentice. The two facets of the apprentice-firm relationship give rise to two components of the financial relationship between apprentices and firms. In reality, however, the monetary arrangements surrounding apprenticeship training in Ghanaian manufacturing are more complex. In addition to wage payments, many apprentices receive allowances for food, housing, clothing, or transportation. These payments are made both in cash and in kind but are identified by firms as distinct from wage payments. In larger firms, these allowances are paid to all workers and have been used as a way to generate nontaxable compensation.1 In smaller firms, allowances are often paid to apprentices even when wages are not. Historically, apprentices often lived with their masters during training. Although this practice has largely died out, provision of basic living expenses by the master continues to be a common feature of these contracts." [velenchikApprenticeshipContractsSmall1995]** -->

\FloatBarrier 

# Data and Methods {#data}

## Country Context

Despite the relative stability of its democratic government and strategic importance as a transportation hub for West Africa, Bénin (population approx. 12.1 million) performs poorly on many development indicators, ranking 158th out of 189 countries on the 2020 Human Development Index. Youth employment is a particularly pressing issue, with youth labor force participation decreasing from 53.9 percent in 2003 to 32.8 in 2018\footnote{According to the National Household Survey, Enquête de Suivi de l'Enquête Modulaire et Intégrée sur les Conditions de Vie des Ménages (HIES). The labor force is the sum of all persons in a given age range who are employed and those who are unemployed. The youth labor force participation rate is the total size of the labor force aged 15-24, as a percent of the population in the same age range.}. As in other parts of SSA, secondary and tertiary school enrollment has seen a steady increase in the past two decades, coinciding with a sharp decrease in the youth employment rate: according to the most recent labor force surveys, the youth employment-to-population ratio decreased by 20 percent, from 40 to 31.5 percent, between 2010 and 2018, compared to an 13 percent decrease for adults age 25 or older over the same time period\footnote{The employment-to-population ratio is the proportion of a country's population in a given age range that is employed. Youth are defined to be aged 15-24.}. Meanwhile, the share of youth neither in employment, education or training (NEET) increased from 17 percent in 2011 to 35 percent in 2018 (see Figure \@ref(fig:fig-enrollment)) --- one the highest rates in West Africa, and the world [@ilo2022].

```{r fig-enrollment, fig.cap="Rates of youth enrollment and inactivity: Bénin and SSA", fig.height=4, fig.width=8, , fig.pos='H'}
```

As enrollment in formal education has not translated to increasing rates of youth employment, interest in promoting alternative pathways to the labor force has grown. In Bénin, entry into formal technical and vocational education and training (TVET) begins after the completion of the second year of secondary school, or nine years of education. Yet across the country, the median number of years spent in the education system is just four; only five percent of youth of secondary school age are enrolled in TVET [@ilo2021], in line with the six percent of young workers estimated to participate in formal TVET across SSA [@hofmann2022]. Thus, rather than formal TVET, it is informal apprenticeship that is the primary conduit into the labor market for early school leavers in Bénin, with as many as 300,000 young men and women estimated to be in training [@ilo2021]. Recent examples of investment in Bénin's apprenticeship system include \$6.3 million from the World Bank's for the Benin Youth Employment Project (PEJ), completed in 2019, and a planned \$16.4 million dollar investment in strengthening the TVET system starting in 2020 [@worldbank2020c].

In 2005, the government of Bénin announced a restructuring of traditional apprenticeship in the informal sector. Two national apprenticeship schemes were introduced: a formalization of informal, firm-based apprenticeship called the _Certificat de Qualification aux Métiers_, which introduced a national certificate for the completion of traditional training, and the dual system _Certificat de Qualification Professionnelle_ (CQP) program, which sought to combine in-firm, firm-specific training with more general classroom-based teaching, and to accredit the training though a separate nationally-recognized certificate. The three stated objectives of the CQP reform were to (i) offer practical and theoretical training to youth under apprenticeship contracts in the craft sector (ii) train a high-performance labor force; and (iii) improve the productivity and profitability of workshops in the craft sector [@davodoun2011].

The CQP is currently available for 13 out of the more than 300 trades listed in the craft sector: auto mechanics, motorcycle mechanics, air conditioning installers, tailors, masons, carpenters, metalworkers (primarily welding of gates for living compounds), electricians, and plumbers\footnote{This selection of trades was based at least in part on existing trades from early experimental dual training programs to take advantage of existing training center infrastructure. The CQM is available for about 50 trades.}. To participate in the CQP, applicants must (i) be at least 14 years old, unless otherwise authorized by the labor inspector; (ii) have a written apprenticeship contract that complies with labor laws; (iii) have completed at least 6 years of formal schooling; (iv) pass a national entry examination; and (v) receive funding from the Fonds de Développement de la Formation Professionnelle Continue et de l’Apprentissage (FODEFCA), the government body responsible for procuring and distributing funding for the program [@kofswisseconomicinstitute2017]. Firm owners apply on behalf of the apprentices already in their charge, generally through local craftsmen associations. After three to four years of training, CQP apprentices may attempt the final examination, which takes place on one date a year for the entire nation, has a practical and a written component, and is overseen by state representatives and local craftsmen. Upon successful completion of this exam, apprentices receive a nationally-recognized certificate.

The main government organs tasked with the administration of the CQP are the national TVET directorate (DETFP), which is in charge of apprentice recruitment and training center accreditation, the Direction of Test and Exam Services (DEC), in charge of the entrance and exit examinations, and FODEFCA [@nouatin2019]. The CQP began curriculum planning in 2005 with technical assistance from the French Development Agency (AFD) and the Swiss Agency for Development and Cooperation (SDC), among others, and became operational in 2008. In 2012, management of the program passed from Swisscontact, a Swiss NGO, entirely into the hands of FODEFCA [@nouatin2019]. Cost sharing for the CQP program is shared by the state and the apprentice, with the state financing body for dual apprenticeship, FODEFCA, officially taking on 90 percent of the training costs [@kofswisseconomicinstitute2017]. However, FODEFCA is largely reliant on external donor funding, and regular financing has been an issue for the program in recent years [@david-gnahoui2017]. The financing of dual training comprises three main budget items: the costs of training to the firm, the administration of the vocational training center, and expenses for entry tests, the final examination, and certification. While on-the-job training in the firm is paid for by the apprentice or their parents, activities in the vocational training centers are largely financed through FODEFCA from various sources (national budget, donors, NGOs, etc.). Certification upon successful completion of the CQP is allocated to the national budget via the Directorate of Examinations, DEC [@david-gnahoui2017].

<!---

_From an economic point of view, it seems obvious that a firm would hire the more apprentices the higher the net benefit of apprenticeship training amounts to. So far, most studies suggested that apprenticeship training results in net cost during the apprenticeship (e.g. Beicht et al. 2004). This would imply that companies need to be able to recoup these net cost by employing graduated apprentices as skilled workers and paying them a wage lower than the value of their marginal productivity, because otherwise they would not o↵er an apprenticeship programme. Themostrecent study of the Swiss apprenticeship system shows, unlike other studies, that about two thirds of the training firms find it profitable to train apprentices (Schweri et al. 2003)._

<!-- With the goal of expanding the number of trades eligible for the CQP and CQM, the Béninese government, through the Youth Employment Project (PEJ) co-financed by the World Bank, has focused on expanding the programs to the crafts, tourism and food processing sectors in recent years.

Officially, the CQP recognizes "professional, technological and general skills acquired by the apprentice for the exercise of a professional activity" [@benin2010].

Criticisms of the CQP include the lack of certified and accredited trainers at the training centers and unstable financing, which leads to high fluctuation in the number of applicants admitted to the program [@david-gnahoui2017; @ilo2020a].

*Dual apprenticeship* is a particular form of this institution, most popular in Germany and Switzerland but practiced across Europe, which combines on-the-job training with continued classroom-based general and trade-specific education. Generally, apprentices in the dual system spend between 20 and 30 percent of their training in the classroom, and train for two to four (most commonly three) years, depending on country and trade.  _Theory training enables young people to acquire general education as well as technical and vocational skills._ Due to the additional costs of classroom training, investments per apprentice will be higher, whether these are covered by the apprentice (in the form of a wage reduction), the firm, or state subsidies. Evidence from Germany suggests that dual training results in net costs in the time of training, which is recouped by firms when apprentices stay with the firm after training [@acemoglu1998, @pfeifer2009]; though recent results suggest that net benefits during the training period are closer to neutral when the impact of hiring an apprentice on profits is compared to the hiring of similarly unskilled or semi-skilled workers on the labor market [@mohrenweiser2009]. 

-->

## Sampling and Attrition

The data for this study was collected in two separate surveys. The first consisted of interviews with apprentices who had applied to the 2019 cohort of the CQP program, who took the entrance exam on the 5th of November, 2018. The second survey was conducted with the firm owners, or master craftsmen, of the apprentices' respective training firms. To allow for trade-level controls and to reduce travel distances for interviews, apprentices were randomly selected from a subsample of all CQP applicants: those training in electrical installation, carpentry, masonry, metalwork or plumbing workshops, and training in the southernmost regions of Bénin.

| CQP Status | Explanation | Apprentice Survey | Firm Survey |
|-----------------|-----------------|-----------------|-----------------|
| **Selected** into CQP | Applied to CQP in 2019 through master craftsmen. Passed exam and was admitted to program. | Random sampling from list of all CQP applicants in five chosen trades and from southern Bénin) | Master craftsman assesses *at most* two apprentices, randomly chosen from all CQP selected/not selected in firm at baseline. |
|   |
| CQP applicant but **Not Selected** | Applied to CQP in 2019 through master craftsmen. Was not admitted due to exam score or lack of proximate training center. | Random sampling from list of all CQP applicants in five chosen trades and from southern Bénin) | Master craftsman  assesses *at most* two apprentices, randomly chosen from all CQP selected/not selected in firm at baseline. |
|   |
| **Did Not Apply** to CQP | Did not apply to CQP. Training as traditional apprentice. |  N/A | Master craftsman lists *up to* 5 apprentices who did not apply to CQP. Assesses only one, randomly chosen at baseline.|

Table: (\#tab:sampling) Apprentice Sampling

\noindent In addition to questions regarding training practices and firm performance, master craftsmen were asked to assess specific apprentices training at their firm. Up to two CQP applicants per firm\footnote{Only CQP applicants who had participated in the apprentice survey could be assessed individually by MCs.} were selected for this exercise. In addition, a single apprentice who had not applied to the CQP was assessed by the craftsman in a similar manner, if there was such an apprentice training in the firm\footnote{As the apprentice survey consisted only of applicants to the program, the only data on non-applicants comes from master craftsmen.}. The sampling procedure for choosing the apprentices for these personal assessments is summarized in Table \@ref(tab:sampling) above.

The baseline wave for the two surveys was collected in July-August 2019. The apprentice survey included questions on training characteristics, employment outcomes, their own skills and competences, and the perceived quality of their training, while the firm survey included questions on worker characteristics, wages, and firm expenses and revenues. We additionally surveyed all MCs about the firm's training practices and expenses, as well as their perception of up to three individual apprentices’ skills, experience, diligence, efficiency, learning ability, and so on (sampling procedure detailed above). Data on `r (df %>% group_by(wave) %>% summarise(n_distinct(IDYouth)))[1,2]` apprentices working for `r (df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[1,2]` unique firms was collected at baseline. Descriptive statistics for apprentices and firms are shown in Table \@ref(tab:tbl-cqpdesc) below. 

```{r, tbl-cqpdesc, echo=FALSE}
```

\noindent Summary statistics from the baseline survey show the sample to be predominantly male youth who, though of average age for an apprenticeship at `r round(mean(baseline$age, na.rm = T), 1)` years [@ilo2022], are significantly more educated than is typical for traditional apprentices, with over half having completed at least some secondary schooling. Successful applicants to the CQP have higher education attainment than unsuccessful applicants; all applicants have higher attainment than non-applicants. Applicants also have more experience coming into the CQP than the required six months, with the average applicant having trained for `r mean((df %>% filter(SELECTED == 3))$baseline_duration, na.rm = T)` years at the time of application. Non-applicants are younger, less educated, and less experienced than applicants at baseline, suggesting that master craftsmen send their most able and veteran apprentices to stand for the dual training entrance exam. This may indicate that the CQP is perceived by MCs as something more akin to a continuing education rather than an entry-level apprenticeship program.

The majority of training firms are small workshops comprising the firm owner --- the master trainer --- and several apprentices. Two firm sizes are shown in Table \@ref(tab:tbl-cqpdesc): those stated directly by the firm owner in response to the question, "How many people (including you and your apprentices) are currently working in your business?" and those calculated by the author by summing the number of apprentices, partners, paid and unpaid family workers, and occasional workers reportedly participating in firm activities. Using self-reported size, `r sum(prop.table(table(baseline$firm_size_bins_reported))[1:2])*100`% of firms employed a total of five people or less (including the owner) at baseline and `r sum(prop.table(table(baseline$firm_size_bins_reported))[1:3])*100` percent of firms employed no more than ten (`r sum(prop.table(table(baseline$firm_size_bins))[1:2])*100`% and `r sum(prop.table(table(baseline$firm_size_bins))[1:3])*100` percent, respectively, using author-calculated size). Thus, training firms in the sample are small, in line with observations from the informal sector in Ghana [@frazer2006; @velenchik1995], but also roughly similar to the size distribution of a sample of Swiss firms reported by @muhlemann2007\footnote{Our sample has a higher density of firms with three to nine employees, whereas the Swiss sample (which includes non-training firms) has a higher density of firms with three total employees or fewer. This shift can be explained by the large number of apprentices in the Béninese training firm sample: at baseline, the average firm employed about four apprentices for every other type of employee, or a total of `r round(mean(baseline$FS6.1, na.rm = T), 1)` apprentices.}.

The endline survey was conducted in August-September 2021. Overall apprentice attrition in our sample, at `r round((1-(df %>% group_by(wave) %>% summarise(n_distinct(IDYouth)))[2,2]/(df %>% group_by(wave) %>% summarise(n_distinct(IDYouth)))[1,2])*100,1)`%, is clearly very high, even when compared to studies in similar contexts: for instance, @crepon2019 and @hardy2019 both report youth attrition of around 10%, though it is generally common for studies of training programs to be affected by high rates of attrition [@mckenzie2017]. Both youth non-response and firm non-response\footnote{Data on apprentices who had not applied to the CQP was only obtained from firm owners, and is thus only subject to firm attrition.} drive attrition in our sample. We assess the severity of this problem by testing for differences between retained and attritted apprentices in terms of relevant baseline characteristics such as CQP participation, trade, or socioeconomic variables. Tables \@ref(tab:tbl-attritionapps) in the Appendix gives no indication of such differences. In Table \@ref(tab:tbl-attritionappsreg), also in the Appendix, we estimate a logit model where attrition is a function of CQP status at baseline and other apprentice characteristics; again, we find no systematic bias. At only 1 percent, reported program dropout among CQP participants retained in the sample is very low compared to similar studies, e.g. @crepon2019 who report 31.2% dropout for dual apprenticeships and 32.5% for traditional apprenticeships. Sample attrition may be correlated with program dropout, but we are unable to test for this possibility with the available data. Finally, despite starting with comparable experience at baseline, unsuccessful applicants to the CQP are more likely to have graduated after three years than CQP participants --- Covid-19 related interruptions to training and the single annual CQP examination date both likely contributed to the low graduation rate of CQP apprentices after three years.

Similarly, of `r (df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[1,2]` firms interviewed at baseline, only `r (df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[2,2]` could be contacted at endline, for an overall firm attrition rate of `r round((1-(df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[2,2]/(df %>% group_by(wave) %>% summarise(n_distinct(FS1.2)))[1,2])*100, 1)`%. Tables \@ref(tab:tbl-attritionfirms) and \@ref(tab:tbl-attritionfirmsreg) in the Appendix likewise suggest that firm attrition was not correlated with key firm characteristics such as size or profitability, though a logit model of attrition shows that firms with a higher proportion of CQP participants among their apprentices are somewhat more likely to exit the sample.

<!-- (`r mean((firms %>% filter(wave == 0, firm_size > 1))$FS6.1/((firms %>% filter(wave == 0, firm_size > 1))$firm_size-1), na.rm = T)*100`% of calculated firm size at baseline and `r mean((firms %>% filter(wave == 1, firm_size > 1))$FS6.1/((firms %>% filter(wave == 1, firm_size > 1))$firm_size-1), na.rm = T)*100`% at endline). -->

<!-- 

Though dual training is predicated on classroom teaching about once a week, we found that external training (classes or training that took place outside of their master's workshop) was not limited to CQP participants. At endline, `r mean(as.numeric(endline$YS4.43), na.rm = T)*100`% of apprentices reported participating in such training in the preceding three months. However, only `r mean(as.numeric((endline %>% filter(SELECTED == 1))$YS4.43), na.rm = T)*100`% of CQP participants reported doing so, despite external training being a constituent and necessary component of the CQP program's dual training structure.

By combining time trained at the time of the survey with the expected graduation date, we find that apprentices expect their training to last `r mean(df$projected_duration, na.rm = T)` years, with a standard deviation of `r sd(df$projected_duration, na.rm = T)` years, in line with our assumption of four-year apprenticeships. Metalworking apprenticeships reportedly last almost a year longer than the other trades in the sample.

Apprentices also explained their motivation for starting an apprenticeship. The most frequently cited motive was interest in the trade, followed by the promise of good earning opportunities and the insistence of the parents. They were also asked why they had chosen the craftsmen they were training with; the majority replied that they were attracted by the patron's reputation, or that their parents had made the choice in their stead. As found in previous studies of informal apprenticeship in SSA, most apprentices express a wish to start their own business after graduation (`r sum(df$YS4.53[!is.na(df$YS4.53)]==2)/length(df$YS4.53[!is.na(df$YS4.53)])*100`%) followed by seeking employment with a different workshop (`r sum(df$YS4.53[!is.na(df$YS4.53)]==1)/length(df$YS4.53[!is.na(df$YS4.53)])*100`%). 

Apprentices report working `r round(mean(df$a_weekly_hours, na.rm = T),2)` hours in the previous week, compared to `r round(df %>% group_by(FS1.2) %>% summarise(x = mean(firm_weekly_hours)) %>% select(x) %>% colMeans(na.rm = T),2)` hours reported by the patron.

`r sum(endline$grad == "Graduated", na.rm = T)/nrow(endline)*100`% had graduated after three years, and the majority, `r sum(endline$grad == "Still training", na.rm = T)/nrow(endline)*100`%, were still training. Graduation rates are highest for electrician apprentices, at `r sum((endline %>% filter(FS1.11 == 5))$grad == "Graduated", na.rm = T)/nrow((endline %>% filter(FS1.11 == 5)))*100`%, but comparable across trades.

However, dropout rates suggest that survey attrition is not caused predominantly by apprenticeship leavers: even when considering as dropouts apprentices whose reason for discontinuing training is unknown, just `r sum(endline$grad == "Dropped out" | endline$grad == "Unknown", na.rm = T)/nrow(endline)*100`% of all apprentices for whom data was collected at endline left their program before graduation. -->

## Estimating Apprentice Benefits {#appmethod}

We first examine the benefits accruing to apprentices over the observed time period of three years. These benefits can be separated into two categories: human capital gains and material benefits.

Human capital gains are measured using a set of trade-specific scores measured separately for each apprentice at baseline and endline. These amount to a simplified version of the "task approach" utilized in the technological change literature [see @dicarlo2016; @crepon2019]. Unlike the general tasks used to measure skills in the task approach, however, we measure apprentice knowledge by means of a short test based on CQP curricula. Each question was a multiple choice question, and between 4 and 5 knowledge questions were posed to each apprentice; because apprentices who did not apply to the CQP were not interviewed directly, the knowledge score was only measured for CQP applicants. The questions are reproduced in Appendix \@ref(cqp-appendix-b).

Our measures of competence and experience, on the other hand, are based on a short roster of tasks drafted in collaboration with local craft experts and practitioners, an approach similar to that employed by @hardy2019 (all tasks are shown in Appendix \@ref(cqp-appendix-b)). Firms were asked to assess apprentices on this series of 10 to 15 trade-relevant tasks\footnote{Apprentices were asked to self-evaluate their competence at endline using the same metric. Self-evaluation was not initially planned and thus unavailable at baseline.}, and the percentage of tasks in which apprentices are deemed competent or experienced (on a binary scale) constitute their score in each of the two measures. Similar to the task approach, this method allows for worker-level measurement of ability and experience based on tasks performed. Each apprentice received a score in each of the three dimensions. 

Regression analysis is then used to examine the impact of dual training on these three measures of apprentice learning. We use the specification

$$ y_{it} = a+\sum_{j}\text{status}_{ij}+\sum_{k}\text{status}_{ik}\text{×}Endline_t+{Endline}_t+\mathbf{X}_{it}+\mathbf{Z}_{jt}+u_{it} $$ 
where $y_{it}$ is the outcome for apprentice $i$ at time $t$, $\text{status}_{ij}$ corresponds to apprentice status $j$ of apprentice $i$ in the context of the CQP program: either successful applicant, unsuccessful applicant, or non-applicant, and $\sum_{k}\text{status}_{ik}\text{×}Endline_t+{Endline}_t$ are the interaction of a CQP participation dummy and survey wave and a dummy for CQP non-applicants and survey wave. These interactions identify any gains in learning outcomes that can be linked to CQP participation. $\mathbf{X}_{it}$ is a column vector of apprentice characteristics, $\mathbf{Z}_{jt}$ is a column vector of training-related firm characteristics, $a$ is a constant, and $u_{it}$ is an error term.

Finally, we assess the material benefits of training accruing to the apprentice, calculated simply as total fees paid less allowances received. These two categories of transactions between MC and apprentices are described in greater detail in the next section.

<!-- Fees are typically paid by the parents directly to the MC at the beginning and end of an apprenticeship. Allowances are transfers from the MC to the apprentices and cover general expenses, in which case they are often referred to as "pocket money" or "chop money", as well as specific expenses such as meals or transportation costs.  -->

## Estimating Firm Benefits {#firmmethod}

\input{tables/tbl-cbmodels}

Firm benefits were calculated using the accounting approach [@gambin2013; @muhlemann2014], which relies on survey data from training firms to identify and quantify the costs and benefits arising from training provision. It has only recently started being applied in lower-middle income countries [@renold2018; @bolli2020; @bolli2021]. The accounting approach subtracts the costs of training from its benefits and is a simple way of estimating the value that training apprentices generates for the training firm. In a second step, we use regression analysis to assess the impact of hiring additional apprentices, both traditional and CQP dual trainees, on firm outcomes.

The components of the accounting approach used in this study are summarized in Table \@ref(tab:tbl-cbmodels). Firms receive two primary benefits from training apprentices: apprenticeship fees and the apprentices' productive contributions to the firm. Training fees can be paid in full before the commencement of training or split into payment at the beginning, during, and at the conclusion of training [@velenchik1995]. Five categories of fees were reported by both apprentices and MCs: entry fees, formation (or general training) fees, liberation (or graduation) fees paid at the conclusion of training, fees as compensation for the materials and equipment used in training, contract fees, and application fees\footnote{Fees are often paid in kind rather than in cash.}. Fees were reported as the total paid for the entirety of the apprenticeship; we assume four-year apprenticeships to estimate annual ammounts.

The second benefit of training for firms, apprentices' net productive contributions, were not reported explicitly by the firm owners and thus needed to be estimated with the aid of several assumptions. First, we assume the competitive model of labor markets (with heterogeneous wages), in which workers are paid the marginal product of their labor. Second, we assume apprentice productivity is equal to that of an untrained employee with no more than a primary education for the first two years of training, and increases to that of trained employee for the final two years\footnote{This is a simplification of the approach used by @bolli2020, in which apprentice productivity is estimated to increase linearly from that of an unskilled worker to that of a skilled worker between defined points in their training. A popular alternative to this approach involves eliciting specific tasks performed by apprentices and estimating costs savings based on the wages paid to workers who would otherwise be responsible for said tasks [@hauschildt2018]. Our firm-apprentice data did not cover specific tasks and is thus not equipped to carry out such an analysis.}, and we use detailed wage information reported by firm owners to estimate the average annual productivity over the course of a four-year apprenticeship. Under these assumptions, the annual productive value generated by apprentice work amounts to the average of these two wages.

Costs of apprenticeship for the firm are categorized into three categories: allowances, training expenses, and lost trainer productivity. Allowances are disbursed irregularly by the firm owner for small expenses such as travel and meals. These are reported by firms at the apprentice level (separate reported allowances for each apprentice). To estimate total annual allowance expenditures per apprentice, we thus sum over all allowance categories and assume that apprentices work 20 days per month; the extrapolated monthly sum is then multiplied by the number of months the training firm was operating in the past year to arrive at an annual estimate for each apprentice. Alternative estimates using different workload assumptions or substituting firm-reported with apprentice-reported figures are shown in Table \@ref(tab:tbl-allowancebounds) and Table \@ref(tab:tbl-allowboundsapp) in the Appendix.

To help estimate the training costs paid by the firm, owners were asked to identify any costs directly or indirectly related to their training activities. Specific training costs include equipment costs, which comprise all costs for physical infrastructure necessary for training; raw materials such as cement, lumber, or scrap metal used for demonstration or practice purposes; training equipment such as workbenches, toolkits, or other machines purchased or rented specifically for apprentices; rent for training facilities if training was not conducted exclusively in the MC's workshop; and books and any other training materials. Firms reported the training costs in each category for the month prior to the interview; to estimate annual training costs per apprentice per year, the reported firm-level costs are divided by the number of apprentices training in the firm and multiplied by the number of months the firm was open in the previous year.

An additional cost paid by the training firm comes in the form of time lost by training staff when they would have otherwise been engaged in productive activities. Lost trainer productivity is, like apprentice productivity, estimated using wage data. Hourly wages for skilled employees were calculated from monthly wage data and multiplied by the number of hours that the workshop stopped all productive activities to train apprentices in the previous week, as reported by MC. This estimate is burdened by the largest number of assumptions: it is uncertain whether all employees who train apprentices in the firm (a number reported by the MC) stop work entirely while the workshop takes a break to train; whether the majority of lost productivity occurs during these breaks, or in the otherwise normal operation of the firm during which they must also tend to the apprentices. Moreover, the total duration of these breaks in the past week is a very small sample from which to extrapolate to annual costs. Lacking a better method, we report these estimates as the final cost component. 

<!-- Firms estimated the hours trained on the last day the firm stopped all productive activities to train apprentices, as well as the number of days per week that such training occurred. We use this information as the basis for our estimation. We extrapolate weekly hours of training (hours trained on previous day of training x days trained per week) to annual hours by assuming four work weeks per month and multiplying by the number of months the firm owner reported being open in the previous year. As when estimating apprentice productivity, we set reported monthly wages for skilled workers (wage employees who had trained with the current firm) equal to trainer productivity and divide by the approximate hours worked in the past month to arrive at approximate hourly wage per trainer (assuming four work weeks per month and using firm-reported days open last week and hours worked on the last day). Finally, we multiply by the number of trainers and divide by the number apprentices per firm to arrive at an estimated cost, per apprentice, in terms of total foregone employee productivity resulting from training activities. -->

Net benefits were calculated by subtracting some or all of the costs listed above from some or all of the benefits. The simplest model, Model I in \@ref(tab:tbl-cbmodels) above, reduces the value of training accruing to the firm to the direct material incomes and expenditures from training; namely, apprenticeship fees from the apprentices and their parents, an approximation of training costs, and the total reported allowances disbursed to the apprentice. Models II and III add training expenses and estimated apprentice productivity, respectively, while Model IV combines all four components. Model V includes all costs and benefits, including lost trainer productivity.

<!-- $$ \textbf{Model I:} $$ -->
<!-- $$\textit{annual net benefits}_i = \frac{\textit{fees paid}_i}{4} - \textit{apprentice allowances}_i - \frac{\textit{total training costs}_j}{N_j}$$ 

\noindent for each apprentice $i$ in firm $j$, where $N_j$ is the total number of apprentices in firm $j$.-->

<!-- $$ \textbf{Model V:} $$ -->

<!-- $$ \begin{aligned} \textit{annual net benefits}_i =  &\frac{\textit{fees paid}_i}{4} - \textit{apprentice allowances}_i - \frac{\textit{total training costs}_j}{N_j} + \\ & \textit{estimated apprentice productivity}_i - \\[1em] & \frac{\textit{estimated foregone productivity per trainer}_j}{N_j} \end{aligned} $$

\vspace{.5cm}
\noindent for each apprentice $i$ in firm $j$, where $N_j$ is the total number of apprentices in firm $j$.  -->

In a final step, we study the effect of hiring apprentices, both traditional and those participating in the CQP, on firm size and profits by estimating the following regression:

$$ y_{it} =  a+CQP_i+{Endline}_t+apprentices_{it}+\mathbf{X}_{it}+u_{it}, $$

where $y_{it}$ is the outcome of interest, $CQP_i$ is the number of CQP applicants who were accepted into the 2019 cohort of the program. $apprentices_{it}$ controls for the total number of apprentices training with the firm and not participating in the CQP program (and in contrast to $CQP_i$ is a time-varying measure), while $\mathbf{X}_{it}$ is a matrix of additional covariates for firm $i$ in wave $t$, ${Endline}_t$ is a dummy variable denoting survey wave, $a$ is a constant, and $u_{it}$ is an error term.

\FloatBarrier

# Results {#cqp_results}

##	Impact of Informal and Dual Training on Individuals {#appbenefits}

```{r changes, include = F}
changes <- df  %>% select(IDYouth, comp_all_trades, exp_all_trades, skills_all_trades, wave) %>% pivot_wider(names_from = wave, values_from = c(comp_all_trades, exp_all_trades, skills_all_trades)) %>% mutate(comp_diff = comp_all_trades_1-comp_all_trades_0, exp_diff = exp_all_trades_1-exp_all_trades_0, skills_diff = skills_all_trades_1-skills_all_trades_0)
```

```{r tbl-skills}
```

First, we investigate whether dual training was successful in realizing its primary objective --- increasing the human capital stock of apprentices. To do so, we study the changes in the three human capital indices described in Section \@ref(appmethod) over the observed training period of three years. The changes in the human capital index scores presented in Table \@ref(tab:tbl-skills) indicate that informal apprenticeship training is successful in improving sector-specific human capital of the youth in our sample, both for dual training participants and traditional apprentices: overall competence scores increased by `r mean(changes$comp_diff, na.rm = T)/sd(baseline$comp_all_trades, na.rm = T)` SDs, experience scores by `r mean(changes$exp_diff, na.rm = T)/sd(baseline$exp_all_trades, na.rm = T)` SDs, and knowledge by `r mean(changes$skills_diff, na.rm = T)/sd(baseline$skills_all_trades, na.rm = T)` SDs. Significant improvements in competence and experience are observed for apprentices who participated in, unsuccessfully applied to, and did not apply to the CQP alike, though apprentices who did not apply to the CQP show the largest gains in competence and experience as assessed by their MC. This result is in line with the observation that MCs appear to send youth who are already relatively experienced (and educated) to apply for dual training. 

A paired t-test also indicates significant improvements in competence and experience between baseline and endline across all trades (improvement in competence for plumbing apprentices is marginally insignificant at standard significance levels). On the other hand, gains in the knowledge metric are not statistically significant for any single trade except plumbing - hence, the overall increase in knowledge is driven by results from plumbing apprentices alone. As the average knowledge scores at baseline were significantly lower for plumbing apprentices than for apprentices in other trades, this result may indicate a shortcoming in the metric itself, which was composed of only up to five questions which did not seem to pose a major challenge for most apprentices in the other four trades. Improvement in knowledge was marginally significant (p < 0.10) for CQP participants but not for CQP applicants who were not selected (p < 0.5). 

Although mean human capital accumulation is higher for participating CQP apprentices than non-selected CQP applicants across the three indices, this does not translate into statistically significant differences between the two group (Table \@ref(tab:tbl-skillschangebycqp) in the Appendix). Nor do we observe a significant effect of dual training on the competence and experience indices when apprentices who did not apply to the CQP are added to the control group. To check if apprentices agree with their MC's assessments of their ability, a self-assessment was included in the apprentice survey at endline. Table \@ref(tab:tbl-compexp2) in the Appendix suggests that the two assessments are in general agreement for both the experience and the competence index (the exception again being limited to the plumbing trade, where MCs were relatively critical in their assessment of apprentices' abilities).

<!-- The three metrics used to measure apprentice progress are not comprehensive, in that they do not measure outcomes such as the physical and psychological well-being of apprentices, "soft skills", etc. Thus, we also examine subjective measures of apprenticeship quality captured by a 5-point Likert scale rating of the following characteristics of the firm, as perceived by the apprentice: physical safety of the workshop, the master trainer's knowledge, the master trainer's treatment of apprentices, the treatment of the apprentice by other employees and apprentices, apprenticeship benefits, apprenticeship working hours, the quality of training, and the quality of the equipment and machinery in the workshop. We observe generally positive levels of satisfaction with training quality, ranging from an average of 4 (somewhat good) to 5 (very good) on all categories except physical safety at baseline (average = 3.96) and working hours at endline (3.66). Subjective ratings are also generally constant over time. One exception is satisfaction with working hours, which decreases from an average rating of 4.07 to 3.66 - indeed, we observe an increase in reported working hours from 42 to 45 average hours worked in the past week for apprentices who applied but did not participate in the program (while CQP participants work hours remained stable). Generally, we observe no significant differences in satisfaction between CQP participants and apprentices who applied to the program but did not participate. -->

<!-- To summarize, satisfaction is generally high and stable among all apprentice throughout their training period, though apprentices not participating in the CQP report a sharp increase in working hours towards the tail end of their training. -->


``` {r, tbl-appreg, results = 'asis', fig.align = 'center'}
```

``` {r, tbl-appreg2, results = 'asis', fig.align = 'center'}
```

\noindent The estimated effects of training presented in Table \@ref(tab:tbl-appreg) confirm that apprentice human capital, as measured by the three indices, increases after three years, and that apprentices who do not apply to the CQP program score somewhat lower on the competence index. Higher baseline experience and competence scores for CQP applicants suggest that trainers send their more able apprentices to apply for dual training. The selection process for the program itself, however, does not favor more experienced apprentices according to our metrics (i.e. is as good as random).

The regression estimates suggest that the amount of time the apprentice had trained with the MC prior to the baseline survey, as well as the total number of apprentices training in the same firm as the interviewed apprentice, are correlated with somewhat higher index scores. Participation in the CQP program is shown to have no detectable effect on human capital accumulation, even when the sample is restricted to applicants to the CQP program only (Table \@ref(tab:tbl-appreg2) in the Appendix). 

```{r, tbl-netappbenefits, echo=FALSE}
```

\noindent Table \@ref(tab:tbl-netappbenefits) shows estimates of the apprentices' net cost of training, taking into account training fees and allowances, as outlined in Section \@ref(appmethod). Overall, apprentices receive more in allowances from their trainer over the course of the training period (assuming a four-year training duration) than they (or their parents) pay in total fees. Mean net benefits amount to about `r round((mean(x$a_allow, na.rm = T)-mean(x$a_total_fees, na.rm = T))*605, -3)` FCFA (`r round(mean(x$a_allow, na.rm = T)-mean(x$a_total_fees, na.rm = T), 0)` \$US) per apprentice per year according to apprentices' own estimates and `r round((mean(x$annual_allowances, na.rm = T)-mean(x$total_fees, na.rm = T))*605, -3)` FCFA (`r round(mean(x$annual_allowances, na.rm = T)-mean(x$total_fees, na.rm = T), 0)` \$US) according to the firm survey.

Formation fees (i.e. general training fees) constitute the largest transfer from the apprentice to the MC. Apprentices report significantly higher fees than the MC, for the formation fee in particular. Firms may under-report fees to avoid accusations of gauging, but are at the same time likely to have more direct knowledge of all fees than apprentices, whose parents and relatives usually pay the craftsmen directly. 

According to the figures reported by apprentices, there is no difference between CQP participants and those who applied but were not accepted into the CQP program in terms of fees, allowances, or net benefits. According to MCs, on the other hand, there are large and significant differences in fees paid and allowances received across CQP status, resulting in significant differences in net benefits. Namely, MCs report charging lower fees (in particular formation fees) and distributing higher allowances to CQP applicants (successful and unsuccessful) than to non-applicants. This amounts to `r round((mean((df %>% filter(SELECTED != 3))$annual_allowances, na.rm = T)-mean((df %>% filter(SELECTED != 3))$total_fees/4, na.rm = T)-(mean((df %>% filter(SELECTED == 3))$annual_allowances, na.rm = T)-mean((df %>% filter(SELECTED == 3))$total_fees/4, na.rm = T)))/605,0)` \$US more paid per year by non-applicants per year, on average\footnote{These figures are difficult to explain, as most CQP applicants would have established the terms of their agreement with the MC several years before actually applying to the dual system program, and there is no apparent reason why dual system applicants should be charged less than other apprentices. One possibility is a reporting bias: we have seen that CQP applicants had been training longer at the time of the baseline survey: thus, any fees paid at the onset of training would have been paid earlier by these apprentices than non-CQP apprentices, resulting in possible recall bias on the part of the trainers.}

<!---
Most classroom training takes place in training centers that are less than an hour away from the apprentices' domicile. 

According to survey data, about a quarter of respondents say training does not take place at a training center, but rather at another firm. At endline, when the answer choice was first included, 50% of apprentices reported attending a technical high school and 44 reported attending a training center. 70% report continuing CQP classes during the Coronavirus pandemic.

MCs indicate that about half of all allowances are in the form of general "pocket money".
--> 

## Impact of Informal and Dual Training on Firms  {#firmimpact}

```{r, tbl-netbenefits, echo=FALSE}
```

\noindent Next, we study the net benefits from apprenticeship training accruing to firms. We also investigate whether the firms training more CQP apprentices experience higher productivity growth over the three-year period under investigation.

To first measure benefits in an accounting sense, we use the five "models" summarized in Table \@ref(tab:tbl-cbmodels) and described in detail in Section \@ref(firmmethod). Table \@ref(tab:tbl-netbenefits) above shows the annual costs and benefits estimated per apprentice, group by CQP status. Mean net benefits range from `r round(mean(baseline$cb_I/605, na.rm = T),0)` \$US to `r round(mean(baseline$cb_V/605, na.rm = T),0)` \$US per apprentice per year, depending on the model used. Apprentice allowances and productivity are the decisive factors determining whether training is profitable for a given apprentice; the sign of the mean of each of the net benefit models thus hinges on the inclusion of the apprentice productivity variable (not included in Models I and II; included in Models III-V).

Firm owners report receiving around `r round(mean(baseline$annual_fees, na.rm = T)/605*4, 0)` \$US in total fees per apprentice, or `r round(mean(baseline$annual_fees, na.rm = T)/605, 0)` \$US per year (assuming 4 years of training). This implies a minor increase in the costs of training in Bénin over the past two decades: for instance, @walther2007 report total fees ranging from 50,000 to 150,000 FCFA (96-290 \$US, inflation adjusted). Formation fees represent the largest single fee paid to the firms and account for over half of total fees paid. Other minor fees cover the provision of equipment and materials, application fees (pertinent for the CQP, as the master trainer must submit paperwork in their apprentices' stead), and initiation and graduation fees. As we have seen, apprentices who do not apply to the CQP pay higher training fees and receive less in allowances (both differences significant at the 5 percent level). This also contributes to the significantly lower net costs for this group when estimated with Models I and II.

Apprentice productivity is calculated using firm-level wage information, an approach similar to apprenticeship cost-benefit studies by @wolter2015, @muhlemann2018, @bolli2020, and @bolli2021 and described in Section \@ref(appmethod). The annual productivity estimates reported in Table \@ref(tab:tbl-netbenefits) are an order of magnitude higher than the estimated annual benefits from training fees. Because the wage data necessary for estimating apprentice productivity is not available for many of the small firms in the sample, however, apprentice productivity can only be estimated for about a quarter of the sample.

Total expenses for training amount to `r round(mean((baseline %>% filter(SELECTED == 1))$total_costs, na.rm =T)/605,0)` \$US per year for CQP participants, and `r round(mean((baseline %>% filter(SELECTED != 1))$total_costs, na.rm =T)/605,0)` \$US per year for all others. These estimates suggest a significant increase in the costs of training for CQP apprentices over the past decade: @david-gnahoui2017, citing Zinsou (2012), reported total costs of 100,000 to 250,000 FCFA (\$165-\$413) for a complete CQP training program in 2012, a total that is exceed by reported expenditures for rent, equipment, books, and raw materials alone (for a four-year apprenticeship) in our data.

Allowances paid to apprentices by MCs comprise the largest reported cost of training for firms, accounting for `r  mean(baseline$annual_allowances, na.rm = T)/mean(baseline$total_costs, na.rm = T)*100`% of all training costs. Table \@ref(tab:tbl-netbenefits) shows that allowances disbursed to CQP participants and CQP applicants are significantly higher than to non-applicants; as applicants tend to be older and more experienced apprentices, this may suggest that apprentice productivity (or at least their remuneration) tends to increase over time. Expenditures on equipment, raw materials and rent for training spaces, grouped together as non-wage training costs, are the second-largest expenditure for firms and average `r round(mean(baseline$annual_training_costs, na.rm =T)/605,0)`\$US, while lost trainer productivity only totals about `r round(mean(baseline$annual_foregone_prod, na.rm =T)/605,0)`\$US --- an order of magnitude less than the estimated expenditure on allowances. These various costs of training per apprentice reported by the MCs are summarized in Figure \@ref(fig:fig-costspie) in the Appendix. 

The distributions of per-apprentice net benefits have long left tails for Models I and II and long right tails for Models III-V (plotted with their means in Figure \@ref(fig:fig-apphist) in the Appendix). The left tails are a consequence of the high number of apprentices in a number of firms generating unrealistic annual allowance totals. These are more than compensated by apprentice productivity estimates when included in Models III, IV, and V, skewing the distribution to the right for these models. For models excluding apprentices' productive contributions to the firm (Model I and Model II), we find that apprentices who applied to the CQP program (selected and not selected) are significantly more costly to train (incur higher net costs) on average than non-applicants, on account of the higher allowances MCs report they receive. Only `r round(sum(baseline$cb_II>0, na.rm = T)/sum(!is.na(baseline$cb_II))*100, 0)`-`r round(sum(baseline$cb_I>0, na.rm = T)/sum(!is.na(baseline$cb_I))*100, 0)` percent of apprentices generate positive net benefits when productivity is left unaccounted for (Models I and II). When apprentice productivity is included in the estimate in Models III, IV, and V, `r round(sum(baseline$cb_V>0, na.rm = T)/sum(!is.na(baseline$cb_V))*100, 0)`-`r round(sum(baseline$cb_IV>0, na.rm = T)/sum(!is.na(baseline$cb_IV))*100, 0)` percent of apprentices are estimated to provide a net benefit to their training firm.

Estimates for Models III-V are only available for a small number of apprentices due to the requirement that net benefits only be computed when data for all cost and benefit categories included in the respective model are available. The variance in benefits is high, ranging from net benefits of `r round(max(baseline$cb_III/605, na.rm = T), -2)` \$US to net costs of  `r round(min(baseline$cb_V/605, na.rm = T)*-1, -2)` \$US per apprentice per year. Though large in magnitude, the differences in net benefits between the three CQP groups are not statistically significant due to the limited sample and high variance for Models III and IV. Differences in net benefits are, however, significant for Model V, which includes all cost and benefit categories: CQP apprentices are nearly break-even, while unsuccessful CQP applicants generate benefits of `r mean((baseline %>% filter(SELECTED == 1))$cb_V/605, na.rm = T)` \$US and non-applicants generate net costs of `r mean((baseline %>% filter(SELECTED == 3))$cb_V/605, na.rm = T)` \$US. Unfortunately, net benefits could only be estimated for 28 apprentices using this model. When only unsuccessful CQP applicants are used as the control group, net benefits are significantly lower for CQP participants according to Models IV and V (see Table \@ref(tab:tbl-appnetbenefitsnodna) in the Appendix). Further tabulations can be found in the Appendix: Table \@ref(tab:tbl-appnetbenefitsbywave) compares baseline and endline results, while Table \@ref(tab:tbl-appnetbenefitsbytrade) reports cost and benefit estimates by trade. 


<!-- 
Mean net benefits per apprentice per year range from `r mean((baseline %>% filter(SELECTED == 1))$cb_V/605, na.rm = T)` \$US to `r mean((baseline %>% filter(SELECTED == 1))$cb_III/605, na.rm = T)` \$US for CQP apprentices, `r mean((baseline %>% filter(SELECTED == 0))$cb_V/605, na.rm = T)` \$US to `r mean((baseline %>% filter(SELECTED == 0))$cb_III/605, na.rm = T)` \$US for unsuccessful CQP applicants, and `r mean((baseline %>% filter(SELECTED == 3))$cb_V/605, na.rm = T)` \$US to `r mean((baseline %>% filter(SELECTED == 3))$cb_III/605, na.rm = T)` \$US for non-applicants.
-->


<!-- this rises to about 70\% for Models III-V. According to Model I, `r round(sum(baseline$cb_I>0, na.rm = T)/sum(!is.na(baseline$cb_I))*100, 2)`\% and `r round(sum(endline$cb_I>0, na.rm = T)/sum(!is.na(endline$cb_I))*100, 2)`\% of apprentices are estimated to generate a positive net benefit during training at baseline and endline, respectively; using Model II, `r round(sum(baseline$cb_II>0, na.rm = T)/sum(!is.na(baseline$cb_II))*100, 2)`\% and `r round(sum(endline$cb_II>0, na.rm = T)/sum(!is.na(endline$cb_II))*100, 2)`\% of apprentices are estimated to do so; using Model III, `r round(sum(baseline$cb_III>0, na.rm = T)/sum(!is.na(baseline$cb_III))*100, 2)`\% and `r round(sum(endline$cb_III>0, na.rm = T)/sum(!is.na(endline$cb_III))*100, 2)`\% of apprentices are estimated to do so; using Model IV, `r round(sum(baseline$cb_IV>0, na.rm = T)/sum(!is.na(baseline$cb_IV))*100, 2)`\% and `r round(sum(endline$cb_IV>0, na.rm = T)/sum(!is.na(endline$cb_IV))*100, 2)`\% of apprentices are estimated to do so; using Model V, `r round(sum(baseline$cb_V>0, na.rm = T)/sum(!is.na(baseline$cb_V))*100, 2)`\% and `r round(sum(endline$cb_V>0, na.rm = T)/sum(!is.na(endline$cb_V))*100, 2)`\% of apprentices are estimated to do so. -->

``` {r tbl-cblong, echo = FALSE}
```

\noindent In a small number of firms, differences in net benefits for individual firms results in apprentices with both positive and negative net benefits to appear in the same firm, resulting in an ambiguous net benefit for the firm. Moreover, we would like to determine the size of the total costs and benefits of apprenticeship training --- i.e., for all apprentices in the firm --- relative to total firm revenues and expenditures. To this end, we estimate firm-level net benefits, calculated as the average costs and benefits of all observed apprentices in a given firm multiplied by the total number of apprentices training at that firm\footnote{As training costs are reported at the firm level and apprentice and trainer productivity estimates were based on firm-reported wage schedules and training frequency, we in effect only extrapolate fee and allowance totals.}. Firm-level net benefits of training are then calculated using Models I-V from before, and are shown in Table \@ref(tab:tbl-cblong) above.

Mean estimated benefits per firm average range from `r round(mean(df1$cb_II/605, na.rm = T), -1)` \$US per year to `r round(mean(df1$cb_V/605, na.rm = T), -1)` \$US per year, with the large variance in estimates once again driven by apprentice productivity estimates. Table \@ref(tab:tbl-cblong) reports estimates by firm size, showing that the largest firms in the sample, through significantly higher reported wages, benefit from significantly higher estimated apprentice productivity. Allowances are, as in the individual-level estimations, by far the largest cost related to training. Firm-level aggregation suggests that our methodology may in fact overestimate apprentice allowances: for all but the smallest firms, total estimated allowances are on average higher than total firm expenditures reported by the firm. 

As with the individual-level estimates, the majority of firms are clustered around zero net benefits for all cost-benefit models, with long left and right tails depending on the model used. Models I and II exhibit long left tails, while Models III-V have long right tails, albeit for fewer observations (plotted with their means in Figure \@ref(fig:fig-firmhist) in the Appendix).  According to Models I and II, `r round(sum(baseline1$cb_I>0, na.rm = T, na.rm = T)/sum(!is.na(baseline1$cb_I))*100, 0)` and `r round(sum(baseline1$cb_II>0, na.rm = T)/sum(!is.na(baseline1$cb_II))*100, 0)` percent of firms are estimated to earn a positive net benefit from training (positive net benefits) respectively; when apprentice productivity is included in the estimate, this rises to `r round(sum(baseline1$cb_III>0, na.rm = T)/sum(!is.na(baseline1$cb_III))*100, 0)`, `r round(sum(baseline1$cb_IV>0, na.rm = T)/sum(!is.na(baseline1$cb_IV))*100, 0)`, and `r round(sum(baseline1$cb_V>0, na.rm = T)/sum(!is.na(baseline1$cb_V))*100, 0)` percent, according to Models III, IV, and V, respectively.

<!-- According to Model I, `r round(sum(baseline1$cb_I>0, na.rm = T, na.rm = T)/sum(!is.na(baseline1$cb_I))*100, 2)`\% and `r round(sum(endline1$cb_I>0, na.rm = T)/sum(!is.na(endline1$cb_I))*100, 2)`\% of firms are estimated to earn a positive net benefit from training (positive net benefits) at baseline and endline, respectively; using Model II, `r round(sum(baseline1$cb_II>0, na.rm = T)/sum(!is.na(baseline1$cb_II))*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_II>0, na.rm = T)/sum(!is.na(endline1$cb_II))*100, 2)`\% at endline, using Model III, `r round(sum(baseline1$cb_III>0, na.rm = T)/sum(!is.na(baseline1$cb_III))*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_III>0, na.rm = T)/sum(!is.na(endline1$cb_III))*100, 2)`\% at endline; using Model IV, `r round(sum(baseline1$cb_IV>0, na.rm = T)/sum(!is.na(baseline1$cb_IV))*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_IV>0, na.rm = T)/sum(!is.na(endline1$cb_IV))*100, 2)`\% at endline; using Model V, `r round(sum(baseline1$cb_V>0, na.rm = T)/sum(!is.na(baseline1$cb_V))*100, 2)`\% are estimated to do so at baseline and `r round(sum(endline1$cb_V>0, na.rm = T)/sum(!is.na(endline1$cb_V))*100, 2)`\% at endline. -->


``` {r tbl-firmregs, results = 'asis'}
```


\noindent In addition to direct financial benefits associated with training, which are reflected by a positive balance in the net benefit calculations presented above, apprenticeship training may affect firm productivity through a variety of additional channels. For the CQP program in particular, participating apprentices may acquire skills at a faster pace than their traditional counterparts as a direct result of their theoretical training. Moreover, firms may experience positive spillovers from training, for instance through introduction to a new technology or learning how to better operate machinery already in the workshop. Additional apprentices in general may improve firm productivity by encouraging the owner to hire more employees (e.g. as trainers) or through investments in additional machinery. Indeed, evidence from previous studies indicates that small firms in Uganda and Ghana, when randomly assigned apprentices to train, reported increased profits of up to 15\% per apprentice [@hardy2022; @alfonsi2020]. 

The results from OLS regressions on three firm outcomes are shown in Table \@ref(tab:tbl-firmregs). We find that firm which sent more apprentices to the CQP did not report higher firm revenue or profit growth over the three years under observation. Firms with more non-apprentice workers are found to be more profitable on average after controlling for the number of apprentices training in the firm. Firms with more apprentices, on the other hand, generate higher revenues (after controlling for non-apprentice workers), but these revenues to not translate to higher profitability: a 10 percent increase in the number of apprentices is associated with revenue increases of about 2.5 percent, but not with any significant changes in profits. This suggests that apprentices' productive contributions are often offset in large part by the costs of their training. Finally, we note that reported firm revenues increase by close to 50\% between baseline and endline, but are offset by rising costs and wages, to the point of eliminating any observed growth in profitability. 

# Conclusion {#conclusion}

This study of informal apprenticeship makes several contributions to the literature. First, it is one of the first papers to contrast the efficacy of dual system training in the informal sector to informal apprenticeship. Second, it analyses the benefit of apprenticeship training from the perspective of both the apprentices and the training firms.

Our findings show that, after three years of training, there are no measurable benefits of dual training in terms of human capital accumulation or firm outcomes. Applicants to the CQP as a group progress less than non-applicants, likely because CQP applicants tend to be older, more experienced, and better educated. There are several possible explanations for this disappointing result: most importantly, disruptions to the training schedule as a result of the Covid-19 pandemic caused many days of theoretical teaching to be grouped into longer sessions or to be cancelled outright. However, funding disruptions, poor attendance of theoretical classes, and insufficient trainer qualification at the vocational centers may have also negatively influenced learning outcomes. These and other issues discussion are at the center of an ongoing discussion concerning the history and implementation of the CQP program [@davodoun2011; @david-gnahoui2017; @bankole2020], to which our study contributes valuable empirical evidence.

<!-- an institution which is not only an important source of skills for youth seeking an alternative to the highly competitive formal wage employment, but also a vital source of labor for informal firms --- in the workshops we interviewed, apprentices accounted for xx% percent of the workforce, not counting the firm owners. Finally, we show that, despite not being nominally called "wages", allowances for apprentices constitute the largest expense for firms, and this expense is not defrayed by fees in most cases. Apprentice productive contributions to firm earnings must be substantial to explain the popularity of training among small informal firms and workshops. As few apprentices stay on with the same firm after graduation, the apprenticeship system can be said to more closely approximate the Swiss system than the German system, as apprentices repay MCs via productive contributions during the apprenticeship rather than thereafter. -->


We also find that informal apprenticeship training in general improves apprentice human capital: faced with the same battery of sector-specific knowledge questions after three years, apprentices' scores increased by .13 standard deviations, while their master trainers' assessment of their competence and experience on a series of sector-specific tasks increased by 0.46 SDs and 0.58 SDs, respectively. 

Finally, the results suggest that while informal apprentices in Bénin generally receive no formal wages, small, irregular allowances disbursed to youth for small expenses nevertheless contribute significantly to total firm expenditures. `r round(mean(baseline$annual_allowances>baseline$annual_fees, na.rm = T)*100,0)` percent of apprentices receive more in allowances than they pay in fees when apprenticeships are assumed to last four years. The average apprentice receives `r round(mean(baseline$annual_allowances-baseline$annual_fees, na.rm = T)/605, 0)` \$US more per year in allowances than he or she pays in fees assuming a four year apprenticeship --- a significant proportion of total firm revenues. Considering that the average firm reported training `r round(mean(baseline$FS6.1, na.rm =T),0)` apprentices (`r round(mean(baseline$FS6.1/baseline$firm_size, na.rm =T)*100,1)` percent of all workers in the firm), it becomes clear that apprentices' productive contributions to the firm must be substantial.

Measuring contributions using skilled and unskilled wages and reported hours worked results in very high apprentice productivity estimates, and only for a fraction of firms. These estimates suggest that apprentices contribute about `r round(mean(baseline$annual_app_prod, na.rm =T)/605,-2)` \$US in productive contributions per year. An alternative would be to assume that, due to perfectly competitive markets, firms break even on training. If we assume that allowances and reported monthly training costs are the only expenses related to training, and apprentice fees and productivity the only benefits, firm responses suggest an apprentice productivity equal to`r round(mean(baseline$cb_II, na.rm =T)/-605,-1)` \$US annually. Firms' total net benefits from apprenticeship training are large and positive when these estimated apprentice productivity are included (especially for large firms), and nearly as large but negative when they are excluded. Our contribution is to highlight the relative importance of apprentice productivity and allowances in cost-benefit analyses of informal apprenticeship, though our data does not allow us to pinpoint precise net benefits for all firms.

The first policy recommendation stemming from this research is to encourage master craftsmen to send younger, less experienced youth to apply to the CQP program. First, this is the target group of the program, and second,  theoretical training is likely to be more effective if it coincides with practical training, rather than taking place after most of the practical skills have been attained.

Second, we suggest an increased emphasis on the quality and regularity of classroom training. Ensuring that vocational centers employ certified trainers, providing teacher training, and returning to a weekly schedule for theoretical training.

Third, to encourage regular apprentice attendance and incentivize MCs to release their apprentices to dual training once a week, dispense allowances from the training center. These allowances would be budgeted under FODEFCA's contributions to the CQP. This paper finds that allowances represent the largest cost related to training for firms; a 20 percent reduction in this cost should provide a significant incentive ensuring regular apprentice attendance.

In summary, our results show the tangible and significant skill acquisition that informal apprenticeship can provide. Apprentices in our sample are the beneficiaries of both human capital and financial capital from their master trainers. However, the number of apprentices hired relative to other workers suggests that apprentices repay their firms through substantial productive contributions. This suggests a model of informal apprenticeship that diverges from previous models, which emphasize the role of fees exchanged for specific human capital rather than productive contributions exchanged for allowances.

<!-- @frazer2006 builds a formal theoretical model of informal training where net benefits enter as a single variable. He writes, "apprentices pay fees at the beginning and end of the apprenticeship, but may also receive a small allowance over the course of the apprenticeship. Apprentices also contribute to firm product. The apprenticeship fee, c, should therefore be thought of as the net value of the transfer (pecuniary and non-pecuniary) from the apprentice to the firm owner." This paper argues that the allowances paid are rather considerable in comparison to the fees paid.  -->

<!-- @velenchik1995 argues that training apprentices is a significant source of revenues for firm owners. We find that while this may be so, the source of these revenues is more likely to be the productivity of the apprentices rather than the fees paid. -->

<!-- Policy: -->
<!-- 1. reduce obstacles to applying for the CQP. there is a clear selection into the program, and seems to favor apprentices who are more experienced and educated. perhaps it is seen as a transition to the formal sector. "Both the apprentices and their employers show more interest in the CQM than the CQP, as the conditions for the CQP – namely required educational level (completed primary education), attendance at a (dual) training centre, preparation for and successful completion of an admission test, and the irregular financing of the FODEFCA – constitute obstacles that they cannot overcome. If these conditions are not reviewed or improved, there is a risk that the CQP will lose its candidates" -->
<!-- 2. improve quality of training. lack of quality control has been mentioned in past audits. return to weekly model of training. improved quality will attract more candidates, improve the perception of the certificate, and help justify the costs, mentioned above, that temper master trainer' interest in the CQ. some ideas, such as teacher and MC training, have been suggested already -->
<!-- 3. dispense allowances via training center. cover travel expenses. even if just once a week, our results suggest that it could defray a major cost to trainers. At the same time, it would provide an incentive to apprentice and firm alike to attend training. -->
<!-- 4. provide incentives for trainer participation in dual training. fund the program using apprenticeship tax, but give MCs greater voice in the design of curricula and implementation of training. this can be accomplished by increasing the role of craft associations in the running of training centers, both public and private. -->

<!-- Apprenticeship is not such an important training institution without reason. The number of apprentices shows that it makes sense financially; the high reported allowances relative to fees suggests that apprentices earn their keep. Dual training appears to be perceived as an accreditation-upgrading system for experienced apprentices rather than a full-fledged dual training program. The program should target younger apprentices and focussing on the quality of training in order to ensure the  -->


<!---
This paper analyses the costs and benefits of apprenticeship training with informal firms in Bénin, with a particular focus on the 2019 cohort of apprentices participating in the CQP dual training program. Apprentice-firm data is used to (1) estimate the human capital gains accumulated by apprentices over the three-year observation period, (2) calculate the total benefits (primarily in the form of apprenticeship fees received) and (3) costs of training reported by firms, (4) calculate the net benefits accruing to firms using two models, one estimating apprentice contribution and foregone trainer productivity and one not, and (5) identify the association between apprenticeship training, and participation in the CQP program in particular, on firm profitability, apprentice human capital gains, and the net benefits accruing to firms per apprentice.

First, we find that there is visible selection into the CQP program. CQP applicants are better educated, older, and have spent more time as apprentices than other apprentices in the firm. Firms sending apprentices to the CQP have fewer than 10 total workers on average, of whom the majority are apprentices. 

Human capital gains as measured by trainers' subjective appraisal of apprentice experience and competence were significant, with improvements of .73 and .66 standard deviation between survey waves, respectively. Gains were concentrated in the masonry and carpentry sector, in larger firms, and among apprentices who did not apply for the CQP program.

Two cost benefit models help us understand how net benefits are distributed across firms and apprentice types. When only fees, allowances, and reported training costs are considered, firms suffer a net loss of `r mean(-df$cb_II/605, na.rm = T)` \$US per apprentice per year, and a total of `r mean(-df1$cb_II/605, na.rm = T)` \$US when accounting for the number of apprentices trained. Using reported wages for skilled and unskilled workers to estimate apprentice productive contributions and foregone trainer productivity reduces the net estimated losses, to `r mean(-df$cb_V/605, na.rm = T)` \$US per apprentice per year, and even a net benefit of `r mean(df1$cb_V/605, na.rm = T)` \$US per firm when projecting total benefits at the firm level. Apprentices who do _not_ apply to the CQP program are associated with somewhat lower net losses to their training firms when estimated using the first model.

Regression analysis reveals a stronger relationship between firm size (sans apprentices) and firm profits than exists for apprentice hires (CQP and otherwise). Similarly, apprentices hired at larger firms are associated with higher net benefits. There are no significant differences in effects observed for CQP applicants or CQP participants, though firms appear to select less experienced apprentices to apply for the program.

In sum, we find that training with informal firms fulfills its promise to apprentices, granting them the experience and competence to transition to self-employment upon graduation. Not all firms benefit from training on the balance, with about 60\% of firms suffering net losses from training and training costs generally outpacing the apprenticeship fees paid. Dual training in the form of the CQP program generates few observable benefits, either in terms of apprentice progression or benefits to the training firm. However, the timing of the current CQP cohort, being interrupted in the middle of the training schedule due to the Covid-19 pandemic, may have contributed to lower-than-usual outcomes at both the apprentice and firm level. 


Nevertheless, this paper finds little support for or against the continuation of the program. Given its difficulties with sustainable financing, and in light of criticisms centered on the administration of training centers and qualifications of classroom teachers, helping individual firms defray the high costs of traditional training, and focusing on material and equipment provision for traditional on-the-job apprenticeships, may be a better use of public financing in the short term. For the time being, in any case, dual training only represents a minuscule fraction of all apprenticeship training in the country of Bénin and across West Africa; with the slow growth of formal TVET programs, traditional apprenticeships are likely to remain a promising route for young Africans who lack the financial means or interest to pursue formal education.
_@mohrenweiser2009: The German dual apprenticeship system is often regarded as a potential model for other countries because it allows enterprises to provide and pay for training in specific and general human capital (Harhoff and Kane, 1997; Acemoglu and Pischke, 1999b; Steedman 2001)...We show that employing apprentices in trade, commercial, craft, and construction occupations instead of unskilled or semi-skilled employees has a positive impact on contemporary establishment performance. In contrast, apprentices in manufacturing occupations have higher relative training costs in comparison to their benefits — contemporary establishment performance declines when their share of employees increases while those of unskilled or semi-skilled employees decreases...The net costs assumption stems from a series of descriptive cost–benefit-studies conducted by the BIBB (Bardeleben et al., 1997; Beicht et al., 2004; Wenzelmann et al., 2009). These evaluations calculate relevant costs during apprenticeship training for about 50 occupations and compare them with the economic value of the productivity contribution of a typical apprentice. Evaluations are based on surveys of about 2500 (personnel) managers who assessed one occupation each. They consistently find that all occupations and almost all companies face sizeable net costs during the apprenticeship period...Fougère and Schwerdt (2002) find a positive effect of apprentices on value added only in medium-sized German establishments. Askilden and Nilsen (2005) analyse the recruitment of apprentices during the business cycle in Norway. They find that apprentices are substitutes for skilled workers and are primarily recruited in boom phases. Zwick (2007) finds on the basis of panel estimations that an increase in the share of apprentices does not decrease profits on average_
 
In the German and Swiss context, the primary benefit of training for firms is the productive value of apprentices during and after their apprenticeship. This value is generated when the apprentices is working in the firm, performing skilled or unskilled tasks that would otherwise be carried out by the firm owner, paid employees, or unpaid labor. These benefits accumulate after graduation if the apprentice is hired for less than the labor market wage for workers of equal productivity, as well as any savings on recruitment and hiring costs [@acemoglu1998; @acemoglu1999; @acemogluTrainingImperfectLabour1999].
In comparison to Swiss or German apprentices, youth in SSA are less likely to continue working for their training firm upon graduation. For example, in Ghana, three out of every four apprentices plan to start their own business within five years [@breyer2007; @ahadzie2009]. Many who don't set off on their own are unable to, citing "high start-up costs, the unavailability of land and hostile planning policies tend to obstruct the ability of newly qualified craftsmen to set up on their own" [@ahadzie2009]. Informal firms must thus recoup their costs during the training period (or within a few years) if hiring apprentices is to remain a profitable enterprise. Moreover, workshops rarely hire skilled wage employees, and thus any benefits from "screening" apprentices for their productive capabilities (i.e. wage compression) or savings in hiring costs play a relatively minor role in the training decision (Table \@ref(tab:tbl-cqpdesc)).
If this is the case, why do craftsmen in Bénin train at all? First, apprentices are the cheapest and most abundant source of labor, and can be hired to complete unskilled tasks at little overhead. Second, if trades can be mastered relatively quickly, apprentices will provide skilled labor in the later years of their graduation (and after graduation as well, for as long as they need to work to pay off outstanding training fees or save up for their own workshop). Finally, the apprentices (or, more often, their families) traditionally pay fees to the master craftsmen in return for the provision of training [@velenchik1995; @frazer2006; @bankole2020].

Moreover, recall that the total cost of a four-year CQP training program was estimated to be between \$165 and \$413 in 2012, of which the trainer is covers 10\%. Assuming the most conservative estimate of net benefits (Model II), the average firm owner accrues a net benefit of 38.64 \$US from training a CQP apprentice for four years, about covering the firms 10\% share of the most expensive cost estimate, $413. Assuming Model I and \$413 CQP cost, the training firm nearly covers CQP costs within a single year; when estimated apprentice productivity is included (Model II), firms recover CQP fees nine times over.
Finally, how do these numbers compare to evidence from high-income countries? Due to the disparities in firm size and productivity of informal firms in Bénin and training firms in Germany and Switzerland, where the majority of studies have been conducted, an informative comparison is difficult to make. Though cost-benefit studies from these countries do not report earnings data of training firms, they do suggest that whether firms recoup the costs of training depends on the firm, trade, and region. For instance, @hauschildt2018 reports that in Germany, productive contributions of apprentices covered about 70 percent of a company's training costs, while in Switzerland, net benefits per apprentice per year amounted to about \euro 2,500 [@strupler2012]. Without a clearer understanding of future benefits and better instruments to measure apprentice productivity, the true benefit of apprenticeship training in informal firms will remain ambiguous.

- _The number of trades in which the CQP and CQM are awarded must be increased. Indeed, while the Bénin craft trades register lists 310 trades, the CQP only covers 13 and the CQM fewer than 100. It is essential to develop the design tools needed to organize the CQP and CQM. This requires the training of a sufficient number of methodology experts across all departments nationwide. In addition to these methodology experts, trainers in all crafts and trades need to be trained in sufficient numbers._
- _There are no qualifications that recognize the learning done by employers. As a result, some employers still refuse to enter their apprentices for the CQP or the CQM, as they think that their apprentices will become better skilled than them. In order to maintain the system, it is essential to introduce a mechanism for validating and recognizing the skills of employers._
- _Both the apprentices and their employers show more interest in the CQM than the CQP, as the conditions for the CQP – namely required educational level (completed primary education), attendance at a (dual) training centre, preparation for and successful completion of an admission test, and the irregular financing of the FODEFCA – constitute obstacles that they cannot overcome. If these conditions are not reviewed or improved, there is a risk that the CQP will lose its candidates._
- _Internal and external financial resources must be mobilized to manage the increased number of trades and apprentices involved in the CQP and CQM._
- _The DEC needs to allocate or obtain timely budgetary resources to organize the CQM examinations._
- _UCIM-B demands responsibility for organizing the CQP and CQM examinations in accordance with the West African Economic and Monetary Union (WAEMU) community code on crafts. Dialogue between this body and the ministry responsible for training should facilitate resolving the matter of the transfer of powers_ [@ilo2021]
--->


\clearpage

# References {.unnumbered}
\noindent

<div id="refs"></div>

\clearpage

# Appendix A4 {-#cqp-appendix-a}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A4.\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{A4.\arabic{table}}

```{r, tbl-attritionapps}
```

```{r, attritionappsreg, include = F}
```

```{r, tbl-attritionappsreg, results = 'asis', fig.align = 'center'}
```

```{r, tbl-attritionfirms}
```

```{r, attritionfirmsreg, include = F}
```

```{r, tbl-attritionfirmsreg, results = 'asis', fig.align = 'center'}
```

``` {r, fig-firmsize, fig.width=7,fig.height=4, fig.cap = "Firm size distributions", fig.pos='H'}
```

```{r, tbl-compexp2}
```

```{r, tbl-skillschangebycqp}
```

```{r, tbl-allowances}
```

\renewcommand{\arraystretch}{2}

```{r, tbl-allowancebounds}
```

```{r, tbl-allowboundsapp}
```

\renewcommand{\arraystretch}{1}

``` {r fig-costspie, fig.width=7,fig.height=4, fig.cap = "Breakdown of mean annual training costs per apprentice", fig.pos='H'}
```

``` {r fig-costspie2, fig.width=7,fig.height=4, fig.cap = "Breakdown of mean annual training costs per firm", fig.pos='H'}
```

```{r tbl-wages}
```

```{r, fig-apphist, fig.height=9, out.width="100%", fig.cap = "Distribution of net benefits per apprentice"}
```

```{r, fig-firmhist, fig.height=9, out.width="100%", fig.cap = "Distribution of net benefits per firm"}
```

```{r tbl-appnetbenefitsnodna}
```

```{r tbl-appnetbenefitsbywave}
```

\begin{landscape}

```{r, tbl-appnetbenefitsbytrade}
```

\end{landscape}

```{r, tbl-firmnetbenefitsbywave}
```

\begin{landscape}

```{r, tbl-firmnetbenefitsbytrade}
```

\end{landscape}

```{r, tbl-firmregsfe, results = 'asis'}
```

# Appendix B4 {-#cqp-appendix-b}

\setcounter{figure}{0}
\renewcommand{\thefigure}{B4.\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{B4.\arabic{table}}

|French |English |
|---------------------------------------------------------|---------------------------------------------------------|
| **Masonry** | |
| Lecture d'un plan de construction | Reading a building plan |
| Identification des différents types de briques | Identifying different types of bricks |
| Composition du béton de fondation | Composition of foundation concrete |
| Composition du béton de la dalle | Composition of slab concrete |
| Élévation | Drafting an elevation |
| Chaînage bas | Low trussing |
| Chaînage haut | High trussing |
| Réalisation des pentes | Pouring out inclined surface |
| Pose des hourdis | Laying down slabs |
| Réalisation des poutres | Installing beams |
| Réalisation des feuillures | Installing rabbets |
| Cimentage du plafond | Cementing a ceiling |
| Cimentage du sol | Cementing a floor |
| Pose des chapes | Laying the floorboards |
| Réalisation d'un devis pour une construction | Drawing up an estimate |
|-----------------------------------------------------------|---------------------------------------------------------|
| **Carpentry** | |
| Prise de mesure des portes et fenêtres | Measurement of doors and windows |
| Prise de mesure des tables et chaises | Measurement of tables and chairs |
| Pointage du bois | Scoring of wood |
| Rabotage | Planing |
| Ponçage | Sanding |
| Savoir faire le mastic | Knowing ho to make sealant |
| Assemblage pour la construction d'une chaise | Chair assembly |
| Assemblage pour la construction d'une table | Table assembly |
| Identification des différents bois utilisés | Identification of different woods used |
| Identification des différentes coupures de bois | Identification of different wood cuts |
| Réalisation de devis pour un produit | Drawing up an estimate for a product |
|-----------------------------------------------------------|---------------------------------------------------------|
| **Plumbing** | |
| Lecture d'un plan de plomberie | Reading a plumbing plan |
| Grattage de tuyau | Pipe scraping |
| Collage des raccords | Attachment of fittings |
| Pose des tuyaux | Laying of pipes |
| Réservation des attentes aux poteaux | Securing pipes at the posts |
| Canalisation des tuyaux dans les fausses septiques et puisards | Piping in septic tanks and sumps |
| Canalisation d'un bâtiment | Piping a building |
| Canalisation pour l'alimentation en eau froide | Piping for cold water supply |
| Réalisation d'un devis | Drawing up an estimate |
| Pose apparente des appareils sanitaires | Installation of exposed sanitary appliances |
|-----------------------------------------------------------|---------------------------------------------------------|
| **Metalworking** | |
| Lecture du plan de construction de l'ouvrage | Reading a construction plan |
| Identification des types de feuilles de tôles | Identifying different types of sheet metal |
| Identification des types de barres de fer | Identifying the different types of iron bars |
| Prise de mesure des feuilles de tôles | Measuring the sheet metal |
| Découpage des feuilles de tôles pour la formation de la charpente | Cutting of sheet metal for the frame |
| Prise de mesure pour la formation des fenêtres | Window measurements |
| Prise de mesure pour la formation des portails | Gate measurements |
| Prise de mesure pour la formation des charpentes | Frame measurement |
| Réalisation des cadres pour les fenêtres | Making the frames for the windows |
| Réalisation des cadres pour les portails | Making the frames for the gates |
| Pose des serrures | Fitting the locks |
| Assemblage des feuilles de tôles pour la formation des fenêtres | Assembly of sheet metal for windows |
| Assemblage des feuilles de tôles pour la formation des portails | Assembly of sheet metal for gates |
| Assemblage des feuilles de tôles pour la réalisation des charpentes | Assembly of sheet metal for joining of frames |
| Réalisation d'un devis pour un ouvrage | Drawing up an estimate |
|---------------------------------------------------------|---------------------------------------------------------|
|**Electrical Inst.** | |
| Lecture d'un plan d'électricité | Reading an electrical plan |
| Conception d'un plan d'électricité | Designing an electrical plan |
| Installation du barrage de terre | Installing an earth barrier |
| Tubage du sol | Soil casing |
| Tubage de la dalle | Tubing the floor slab |
| Serrage des boîtiers et coffrets | Clamping of boxes and cabinets |
| Pose des lampes | Installation of lamps |
| Pose des prises | Installation of sockets |
| Installation des disjoncteurs dans les coffrets | Installation of circuit breakers |
| Réalisation d'un devis | Drawing up an estimate |

Table: (\#tab:tasks) Tasks used for assessment of competence and experience

```{r, fig.cap="\\label{fig:metalworkers}Questions for metal workers"}
knitr::include_graphics("figures/questionnaire/metalworking.pdf")
```

```{r, fig.cap="\\label{fig:qs_plumbers}Questions for plumbers"}
knitr::include_graphics("figures/questionnaire/plumbing.pdf")
```

```{r, fig.cap="\\label{fig:qs_carpenters}Questions for carpenters"}
knitr::include_graphics("figures/questionnaire/carpentry.pdf")
```

```{r, fig.cap="\\label{fig:qs_masons}Questions for masons"}
knitr::include_graphics("figures/questionnaire/masonry.pdf")
```

```{r, fig.cap="\\label{fig:qs_electricians}Questions for electricians"}
knitr::include_graphics("figures/questionnaire/electricians.pdf")
```

`r detach_packages(packages)`
