---
title: Benefits and Costs of (Dual) Apprenticeship in an Informal Setting
author: 
  - name: Bart Kudrzycki
    affiliation: Development Economics Group, ETH Zurich, Switzerland, <bartlomiej.kudrzycki@nadel.ethz.ch>
date: \today
abstract: |
   You can include the text for your abstract here. 
thanks: |
  Thanks to Dario for the Markdown template.
keywords: [Keyword A, Keyword B, Keyword, C]
jel: [I24, I32, J15, J16]
bibliography: template-bib.bib
csl: 'apa-no-ampersand.csl'
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    template: template.tex
    extra_dependencies: ["flafter", "placeins"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,
                      warning=FALSE,
                      echo = FALSE, 
                      fig.path='figures/', 
                      floatfoot.envs = c("figure", "table", "longtabu")) 

options(scipen=2, digits=2)
```

```{r loadlibraries, include=FALSE}
library(tidyverse)
library(bookdown)
library(gtsummary)
```

```{r loaddata, include=FALSE}
# load custom functions
source("functions/functions.R")

# load data or run cleaning file in case file does not exist
if (!file.exists("data/cleaned.Rda")) {
  source("code/cleaning.R")
}

# run external analyses
source("code/regression.R")

ggplot2::theme_set(ggplot2::theme_bw())  ## (this just globally sets the ggplot2 theme to theme_bw)
```


# Introduction {#introduction}

\FloatBarrier 

# Background {#background}

\FloatBarrier

# Methodology {#methods}

To estimate the benefits in terms of productivity, we **assume** that apprentice productivity is an increasing step function. Productivity increases with experience. For the first two years, we assume apprentices are as productive as paid family workers

\FloatBarrier

# Data {#data}

\FloatBarrier

# Results {#results}

## Benefits to master trainer

Benefits accrued from productivity include training fees and apprentice productivity.

### Productivity

We first attempt to measure apprentices' marginal productivity. In the competitive model of labor markets (with heterogeneous wages), workers are paid their marginal productivity. We assume competitive labor markets and use wages reported by the master craftsmen for different worker types to estimate apprentice productivity. In other words, worker wages are used as a proxy to differentiate between “skilled” and “unskilled” apprentices productivity. Since apprentices were reported being paid zero wages regardless of experience in almost all cases, we assume zero wages.

Average reported wages are `r mean(wages$app_wages_low, na.rm = T)` FCFA per month for low-skill workers^[Paid family workers, paid workers with no more than primary education only, or occasional workers.] and 11838.71 FCFA per month for high-skill workers^[Paid workers who completed an apprenticeship at the same firm, paid workers who completed an apprenticeship at a different firm, or paid workers with secondary or tertiary education but no apprenticeship.]. Compare to `r mean(wages$FS5.2_1_7, na.rm = T)` FCFA per month that master craftsmen reported paying themselves on average. In sum, the benefits accruing to masters from apprentice productivity can be estimated to be `r 11838.71 - mean(wages$app_wages_low, na.rm = T)` FCFA per month or **`r (11838.71 - mean(wages$app_wages_low, na.rm = T))*12` FCFA** (USD 50) per year.

### Apprenticeship Fees

In contrast to apprenticeship systems found in Germany or Switzerland, fees paid by the families of apprentices are substantial.

```{r summary1}
fees %>% 
  tbl_summary(type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ c("{N_nonmiss}",
                                     "{mean} ({sd})", 
                                     "{median} ({min}, {max})"),
              missing = "no") %>% 
  as_kable_extra(caption = "Summary of fees (in FCFA)", 
                 booktabs = T)
```

`r mean(fees$entry_fee, na.rm = T)` FCFA
`r mean(fees$training_fee, na.rm = T)` FCFA
`r mean(fees$graduation_fee, na.rm = T)` FCFA
`r mean(fees$materials_fee, na.rm = T)` FCFA
`r mean(fees$contract_fee, na.rm = T)` FCFA
`r mean(fees$dossier, na.rm = T)` FCFA

Total fees: `r round(mean(fees$entry_fee, na.rm = T) + mean(fees$training_fee, na.rm = T) + mean(fees$graduation_fee, na.rm = T) + mean(fees$graduation_fee, na.rm = T) + mean(fees$contract_fee, na.rm = T) + mean(fees$dossier_fee, na.rm = T), 1)` FCFA, or USD 250 for approximately four years of apprenticeship.

``` {r fee plot} 
fees_r <- pivot_longer(fees, cols = everything(), names_to = "fee_type") %>% group_by(fee_type) %>% 
  summarise("Mean Fee" = mean(value, na.rm = T))
fee_low <- pivot_longer(fees_low, cols = everything(), names_to = "fee_type") %>% group_by(fee_type) %>% 
  summarise("Fee Low" = mean(value, na.rm = T)) %>% select(`Fee Low`)
fee_high <- pivot_longer(fees_high, cols = everything(), names_to = "fee_type") %>% group_by(fee_type) %>% 
  summarise("Fee High" = mean(value, na.rm = T)) %>% select(`Fee High`)

fees_r <- cbind(fees_r, fee_low, fee_high)

ggplot(fees_r) + geom_bar(aes(x = fee_type, y = `Mean Fee`), stat = "identity", fill="skyblue", alpha=0.8) + geom_errorbar(aes(x = fee_type, ymin = `Fee Low`, ymax = `Fee High`), width=0.2, colour="orange", alpha=0.8, size = 1.2) + coord_flip() + theme_minimal() + labs(x = "", y = "Mean Fee (FCFA)")
```

### Skills

The figures below show changes in skills as reported by the master craftsmen. "Competencies" refer to a set of tasks, unique to each trade, which the master trainer feels the apprentices is capable of carrying out. "Experience", on the other hand, tallies which of these tasks apprentices have done at least once.

Though skills are not directly comparable between trades, masters report their apprentices are capable of completing `r mean(colMeans(competencies, na.rm =  T)*100)`% of trade-relevant tasks at baseline and `r mean(colMeans(end_competencies, na.rm =  T)*100)`% at endline, on average - an increase of `r mean(colMeans(end_competencies, na.rm =  T)/mean(colMeans(competencies, na.rm =  T))-1)*100`%.. By trade, we observe a `r (mean(end_competencies$elec, na.rm =  T)/mean(competencies$elec, na.rm =  T)-1)*100`% increase in reported competence for apprentices in electrical installation, `r (mean(end_competencies$macon, na.rm =  T)/mean(competencies$macon, na.rm =  T)-1)*100`% increase for masonry, `r (mean(end_competencies$menuis, na.rm =  T)/mean(competencies$menuis, na.rm =  T)-1)*100`% increase for carpentry, `r (mean(end_competencies$plomb, na.rm =  T)/mean(competencies$plomb, na.rm =  T)-1)*100`% increase for plumbing, `r (mean(end_competencies$metal, na.rm =  T)/mean(competencies$metal, na.rm =  T)-1)*100`% increase for metalworkers. We thus observe a significant improvement in reported competency.

We can conduct a similar comparison for reported experience: tasks which apprentices have carried out at least once. Masters report their apprentices have experience with `r mean(colMeans(experience, na.rm =  T)*100)`% of trade-relevant tasks at baseline and `r mean(colMeans(end_experience, na.rm =  T)*100)`% at endline, on average - an increase of `r mean(colMeans(end_experience, na.rm =  T)/mean(colMeans(experience, na.rm =  T))-1)*100`%. By trade, we observe a `r (mean(end_experience$elec, na.rm =  T)/mean(experience$elec, na.rm =  T)-1)*100`% increase in reported experience for apprentices in electrical installation, `r (mean(end_experience$macon, na.rm =  T)/mean(experience$macon, na.rm =  T)-1)*100`% increase for masonry, `r (mean(end_experience$menuis, na.rm =  T)/mean(experience$menuis, na.rm =  T)-1)*100`% increase for carpentry, `r (mean(end_experience$plomb, na.rm =  T)/mean(experience$plomb, na.rm =  T)-1)*100`% increase for plumbing, `r (mean(end_experience$metal, na.rm =  T)/mean(experience$metal, na.rm =  T)-1)*100`% increase for metalworkers. We thus observe an even larger increase in apprentice experience over the course of the 3 years between endline and baseline.


``` {r competencies figures}
as.data.frame(cbind(comp = colMeans(end_competencies, na.rm = T), exp = colMeans(end_experience, na.rm = T))) %>% 
  rownames_to_column(., "trade") %>% 
  pivot_longer(cols = c("comp", "exp")) %>% group_by(name, trade) %>% ggplot(aes(x = trade, fill = name)) + geom_bar(aes(y = value),  stat = "identity", position = "dodge") + labs(x = "", y = "% of apprentices", title = "Competence vs. experience at baseline")

as.data.frame(cbind(baseline = colMeans(competencies, na.rm = T), endline = colMeans(end_competencies, na.rm = T))) %>% 
  rownames_to_column(., "round") %>% 
  pivot_longer(cols = c("baseline", "endline")) %>% group_by(name, round) %>% ggplot(aes(x = round, fill = name)) + geom_bar(aes(y = value),  stat = "identity", position = "dodge") + labs(x = "", y = "% of apprentices", title = "Competencies")

as.data.frame(cbind(baseline = colMeans(experience, na.rm = T), endline = colMeans(end_experience, na.rm = T))) %>% 
  rownames_to_column(., "round") %>% 
  pivot_longer(cols = c("baseline", "endline")) %>% group_by(name, round) %>% ggplot(aes(x = round, fill = name)) + geom_bar(aes(y = value),  stat = "identity", position = "dodge") + labs(x = "", y = "% of apprentices", title = "Experience")

```

``` {r competencies by apprenticeship type}

apps %>% 
  mutate(app_type = ifelse(FS7.4 == 3, "CQP", "CQM/Traditional")) %>% 
  select(c("app_type", "elec_score", "macon_score",  "menuis_score", "plomb_score", "metal_score")) %>% 
  tbl_summary(by = app_type,
              type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ c("{N_nonmiss}",
                                     "{mean} ({sd})"),
              missing = "no") %>% 
              add_p() %>% 
  as_kable_extra(caption = "Summary of scores by apprenticeship type", 
                 booktabs = T)

```


# Training Costs

*For  training  companies,  the  most  important  cost  components  are  the  apprentices'  wages,  as  well  as those  of  the  training  staff  (in  both  cases  including  potential  additional  bonuses).  Additionally,  various other  --  mostly  minor  --  expenses  come  into  play:  recruitment,  machinery  and  materials  (as  long  as they  are  only  used  for  training  purposes),  examination  fees,  administration,  and  external  and  internal courses.  Table  1.1  provides  an  overview  of  these  various  cost  positions  (Schweri  et  al.,  2003;  Beicht  et al.,  2004;  Rauner  et  al.,  2009;  Rothbeck,  2016),  which  are  more  or  less  relevant  for  all  companies  that decide to train apprentices. These factors vary according to the different professions trained, according to  branches/sectors,  etc.  (see  section  1.4),  but  can  to  some  extent  be  looked  at  as indispensable/indisputable*.

\FloatBarrier

# Conclusion {#conclusion}

\clearpage

# References {.unnumbered}

::: {#refs}
:::

```{=tex}
\appendix
\clearpage
```
# Appendix Section 

\FloatBarrier
